### 2024-08-30 Tutorial_2_Bayesian/Tutorial_2.Rmd
update Tutorial 2

---
title: "Tutorial_2"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build Bayesian hazard regression models for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent), and visualize the posterior distributions of the effects of congruent and incongruent prime types, relative to blank prime trials.

# Load the libraries that we will be using

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

# Set options.

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()
detectCores()
```

```{r check-info}
packageVersion("cmdstanr")
devtools::session_info("rstan")
```

# Load the person-trial-bin data set that we saved in Tutorial 1.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
ptb_data <- ptb_data %>% filter(period > 5)

# create factor condition, with "blank" as the reference level
ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale; Add dummy variables for each bin.
ptb_data <- ptb_data %>% 
        mutate(period_9 = period - 9,
               trial_c = (trial - 1000)/1000,
               d6  = if_else(period == 6, 1, 0),
               d7  = if_else(period == 7, 1, 0),
               d8  = if_else(period == 8, 1, 0),
               d9  = if_else(period == 9, 1, 0),
               d10 = if_else(period == 10, 1, 0),
               d11 = if_else(period == 11, 1, 0),
               d12 = if_else(period == 12, 1, 0),
               d13 = if_else(period == 13, 1, 0),
               d14 = if_else(period == 14, 1, 0),
               d15 = if_else(period == 15, 1, 0))

head(ptb_data,n=17)
summary(ptb_data)
# 6 subjects
# condition: blank (6401), congruent (2642), incongruent (3797)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Plot the logit and complementary log-log (cloglog) link functions

```{r plot-links}
probability <- (1:99999)/100000
logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
logit    <- function(x) { return( log(x/(1-x)) )}
inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
cloglog         <- function(x) { return( log(-1*log(1-x)) )}

cloglog_prob <- cloglog(probability)
logit_prob <- logit(probability)
dataplot <- cbind(probability,cloglog_prob,logit_prob)

ggplot() +
  geom_hline(yintercept=0, color="white") +
  geom_line(data=dataplot,aes(y=logit_prob,x=probability,colour="logit"),linewidth=1) +
  geom_line(data=dataplot,aes(y=cloglog_prob,x=probability,colour="cloglog"),linewidth=1) +
  scale_color_manual(name = "Link function:", values = c("logit" = "darkblue", "cloglog" = "red")) +
  geom_vline(xintercept = logistic(0), linetype="dotted", linewidth = 0.3) +   
  geom_vline(xintercept = inverse_cloglog(0), linetype="dotted", linewidth = 0.3) +
  annotate("text", x = logistic(0)-.02, y = -6, label = "logistic(0) = 0.5", angle = 90) +
  annotate("text", x = inverse_cloglog(0)+.02, y = -6, label = "inverse_cloglog(0) = 0.6321", angle=90) +
  scale_x_continuous(n.breaks=10, limits = c(0,1), ) +
  labs(x = "Probability",
        y = "logit or cloglog scale") +
  theme(panel.grid = element_blank(),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14),
      #  legend.position="top",
        legend.text = element_text(size=14),
      legend.position = "inside",
        legend.position.inside=c(.2,.75),
      legend.background = element_rect(fill="white") )
```

```{r save-plot-links}
ggsave("Tutorial_2_Bayesian/figures/linkfunctions.png", width = 12, height = 10, dpi = 600)
```

# Visualize different prior distributions on the logit and cloglog scales

To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.

## Simulate, convert, and fit

```{r simulate-convert}
set.seed(11)

logit    <- function(x) { return( log(x/(1-x)) )}
cloglog  <- function(x) { return( log(-1*log(1-x)) )}

# generate draws from U(0,1) and convert
dat <- 
  tibble(p = runif(1e6, 0, 1)) %>% 
  mutate(g = logit(p),
         c = cloglog(p)) 
# display
dat %>%   
  ggplot(aes(x = c)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

plot(density(dat$c))
```

```{r fit-skew-normal}
# fit models
fit_skewN <-
  brm(data = dat,
      family = skew_normal(),
      c ~ 1,
      chains = 4, cores = 4,
      file = "Tutorial_2_Bayesian/models/fit_skewN")
```

```{r}
summary(fit_skewN) 
```

Now we can reverse the process. We simulate from the skew-Normal distribution based on the posterior means for mu, sigma, and alpha, and then convert the results into the probability (i.e., discrete-time hazard) metric.

```{r check-results}
set.seed(11)

inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}

tibble(c = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22) ) %>% 
  mutate(p = inverse_cloglog(c)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

## Visualize how seven prior distributions on the logit and/or cloglog scales look on the probability scale

```{r plot-priors}
logistic <- function(x) { return( 1/(1+exp(-1*x)) )}
logit    <- function(x) { return( log(x/(1-x)) )}
inverse_cloglog <- function(x) { return( 1-(exp(-1*exp(x))) )}
cloglog    <- function(x) { return( log(-1*log(1-x)) )}

set.seed(23)

# A N(0,4) prior on the logit and cloglog scales pushes mass to probabilities of 0 and 1
pr1 <- tibble(prior = rnorm(1e6, mean = 0, sd = 4)) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-13, y=60000, label="N(0,4)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l1 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 4)) %>%  
  mutate(p = logistic(log_odds)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("logistic(prior)") +
  theme(panel.grid = element_blank())

c1 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 4)) %>% 
  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("inverse_cloglog(prior)")+
  theme(panel.grid = element_blank())

# A N(0,2) prior on the logit and cloglog scales pushes mass to probabilities of 0 and/or 1
pr2 <- tibble(prior = rnorm(1e6, mean = 0, sd = 2)) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-6, y=60000, label="N(0,2)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l2 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 2)) %>%  
  mutate(p = logistic(log_odds)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

c2 <- tibble(cloglog_prob = rnorm(1e6, mean = 0, sd = 2)) %>% 
  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

# A student-t(df=7.61) prior with scale 1.57 on the logit scale approximates a uniform distribution on the probability scale. This might be a good prior to use for the alpha parameters or Intercept in a logit-hazard model.
pr3 <- tibble(prior = rt(1e6, df = 7.61)* 1.57) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-12, y=140000, label="t(7.61)*1.57",
              color="red", size=6) +
  annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l3 <- tibble(log_odds = rt(1e6, df = 7.61)* 1.57) %>%  
  mutate(p = logistic(log_odds)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

c3 <- ggplot()

# A Normal(0,1) prior on the logit scale gently regularizes p towards .5. This might be a good prior to use for the beta parameters in a logit-hazard model.
pr4 <- tibble(prior = rnorm(1e6, mean = 0, sd = 1))%>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-3, y=60000, label="N(0,1)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'D', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l4 <- tibble(log_odds = rnorm(1e6, mean = 0, sd = 1))%>%  
  mutate(p = logistic(log_odds)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  geom_vline(xintercept=logistic(0), color="red") +
  theme(panel.grid = element_blank())

c4 <- ggplot()

# A skew_Normal(-0.59,1.26,-4.22) prior on the cloglog scale approxiates a uniform distr. on the hazard scale. This uninformative prior might be good for the alpha parameters or Intercept in a cloglog-hazard model.
pr5 <- tibble(prior = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.22)) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-5.5, y=55000, label="skew_N(-0.59,1.26,-4.22)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'E', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l5 <- ggplot()

c5 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.59, sigma = 1.26, alpha = -4.20)) %>% 
  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

# The skew_Normal(-1,1,-2) on the cloglog scale is a weakly informative prior for the alpha parameters or Intercept in a cloglog-hazard model because hazard values below .5 more likely than values above .5 in general.
pr6 <- tibble(prior = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-5, y=55000, label="skew_N(-1,1,-2)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'F', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l6 <- ggplot()

c6 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-1, sigma = 1, alpha = -2)) %>% 
  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

# A skew_Normal(-0.2,0.71,-2.2) on the cloglog scale gently regularizes p towards .6321. This might be a good prior to use for the beta parameters in a cloglog-hazard model.
pr7 <- tibble(prior = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>%  
  ggplot(aes(x = prior)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(NULL, breaks = NULL) +
  annotate(geom="text", x=-3.2, y=55000, label="skew_N(-0.2,.71,-2.2)",
              color="red", size=6) +
  annotate(geom = 'text', label = 'G', x = -Inf, y = Inf, hjust = 0, vjust = 1, size=8)+
  theme(panel.grid = element_blank())

l7 <- ggplot()

c7 <- tibble(cloglog_prob = rskew_normal(1e6, mu=-0.2, sigma = .71, alpha = -2.2)) %>% 
  mutate(p = inverse_cloglog(cloglog_prob)) %>% 
  ggplot(aes(x = p)) +
  geom_histogram(bins = 50) +
  geom_vline(xintercept=invcloglog(0), color = "red") +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())

(l1 + pr1 + c1) / (l2 + pr2 + c2) / (l3 + pr3 + c3) / (l4 + pr4 + c4)/ (l5 + pr5 + c5) / (l6 + pr6 + c6) / (l7 + pr7 + c7)
```

```{r save-plot-priors}
ggsave("Tutorial_2_Bayesian/figures/plot_of_priors.png", width=14, height=13,dpi=300)
```

# Fit 4 cloglog-hazard models.

## Model M1: general specification of TIME in the baseline condition (blank prime), and main effects of prime type, and trial number.

### Prepare data for M1

```{r data-M1}
# remove unnecessary columns before fitting a model
M1_data <- ptb_data %>% select(-c(bl,tr,trial,period, period_9)) # 12840 obs. of 14 variables
head(M1_data)
summary(M1_data)
```

### Set up priors for M1

```{r priors-cloglog-M1}
priors_M1 <- c(
  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d6"), # weakly informative prior for alpha parameters when using cloglog link
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d7"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d8"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d9"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d10"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d11"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d12"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d13"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d14"), 
  set_prior("skew_normal(-1,1,-2)", class = "b", coef = "d15"), 
  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
)
```

### Fit model M1

```{r fit-model-M1}
plan(multicore)

model_M1 <-
   brm(data = M1_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c +
                          (0 + d6 + d7 + d8 + d9 + d10 + d11 + d12 + d13 + d14 + d15 + condition + trial_c | pid),
       prior = priors_M1,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M1")
```

Model_M1 took about 96 minutes on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).

```{r check-model-M1}
summary(model_M1)
```

## Model M2: third-order polynomical specification of TIME in the baseline condition (blank prime), and main effects of prime types, and trial number.

### Prepare data for M2

```{r data-M2}
# remove unnecessary columns before fitting a model
M2_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
head(M2_data)
```

### Set up priors for M2

```{r priors-cloglog-M2}
priors_M2 <- c(
  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
  set_prior("normal(0, 1)", class = "sd"),                     # prior for standard deviation of random effects
  set_prior("lkj(2)", class = "cor")                           # prior for correlations between random effects
)
```

### Fit model M2

```{r fit-model-M2}
plan(multicore)

model_M2 <-
   brm(data = M2_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c +
                          (1 + period_9 + I(period_9^2) + I(period_9^3) + condition + trial_c | pid),
       prior = priors_M2,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M2")
```

Model_M2 took about .... to run.

```{r check-model-M2}
summary(model_M2)
```

## Model M3: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax proportionality assumption for prime types, and trial number.

### Prepare data for M3

```{r data-M3}
# remove unnecessary columns before fitting a model
M3_data <- ptb_data %>% select(-c(bl,tr,trial,period, d6, d7, d8, d9, d10, d11, d12, d13, d14, d15)) # 12840 obs. of 14 variables
head(M3_data)
```

### Set up priors for M3

```{r priors-cloglog-M3}
priors_M3 <- c(
  set_prior("skew_normal(-0.2,0.71,-2.2)", class = "b"),       # weakly informative prior for beta parameters when using cloglog link
  set_prior("skew_normal(-1,1,-2)", class = "Intercept"),      # weakly informative prior for Intercept when using cloglog link
  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
)
```

### Fit model M3

```{r fit-model-M3}
plan(multicore)

model_M3 <- 
   brm(data = M3_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + # Note that duplicate terms in the model formula are ignored
                           condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),
       prior = priors_M3,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M3")
```

Model_M3 took about .... to run.

```{r check-model-M3}
summary(model_M3)
```

## Model M4: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.

# Compare 4 models

# Plot effects for selected model

##########

Fit a model that relaxes proportionality assumption.

```{r}
# remove unnecessary columns before fitting a model
ptb_data <- ptb_data %>% select(-c(trial_c)) # 12840 obs. of 4 variables
head(ptb_data,n=17)
summary(ptb_data)
```

```{r cubic-random-effects-model}
plan(multicore)

model_cubic_RE <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_cubic_RE")

# the formula above comes down to the following effects:
# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3) ##effect of incongruent interacts cubicly with TIME
```

```{r check-model-cubic-RE}
summary(model_cubic_RE)
```

#### relax all ass

```{r cubictime-quadtrial-model}
plan(multicore)

model_cubictime_quadtrial <-
   brm(data = ptb_data,
       family = binomial(link="cloglog"),
       event | trials(1) ~ 0 + Intercept + 
                           condition*period_9*trial_c + 
                           condition*period_9*I(trial_c^2) + 
                           condition*I(period_9^2)*trial_c + 
                           condition*I(period_9^2)*I(trial_c^2) +
                           condition*I(period_9^3) +
                      (1 + condition*period_9*trial_c +
                           condition*period_9*I(trial_c^2) +
                           condition*I(period_9^2)*trial_c +
                           condition*I(period_9^2)*I(trial_c^2) +
                           condition*I(period_9^3) | pid),
       prior = priors,
       chains = 4, cores = 4, iter = 3000, warmup = 1000,
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 12),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_cubictime_quadtrial")

# the formula above comes down to the following effects:
# cloglog[h(t)]=Intercept+ ##the estimated cloglog[h(t)] in bin 9 for trial 1000 for reference condition "blank"
# period_9+I(period_9^2)+I(period_9^3)+ ##a cubic specification of TIME for reference condition "blank"
# congruent+congruent:period_9+congruent:I(period_9^2)+congruent:I(period_9^3)+ ##effect of congruent interacts cubicly with TIME
# incongruent+incongruent:period_9+incongruent:I(period_9^2)+incongruent:I(period_9^3)+ ##effect of incongruent interacts cubicly with TIME
# trial_c + trial_c:period_9 + ##linear effect of trial number in bin 9 can change linearly over TIME for "blank"
# trial_c:congruent + trial_c:incongruent + ##linear effect of trial number in bin 9 can differ for congruent and incongruent
# trial_c:congruent:period_9 + trial_c:incongruent:period_9 ##how the linear effect of trial number changes linearly over TIME can differ for congruent and incongruent
```

# plot hazard functions

```{r}
model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
model_full_RE_newprior <- readRDS("Tutorial_2_Bayesian/models/model_full_RE_newprior.rds") 
fixef(model_full_RE)
fixef(model_full_RE_newprior)
```

# visualize posterior distributions of the effects of congruent and incongruent primes on hazard relative to blank prime, for each time bin in trial 1000, and for each time bin in trial 100.

```{r load-model}
model_full_RE <- readRDS("Tutorial_2_Bayesian/models/model_full_RE.rds") 
```

```{r make-plot-effects}
post <-as_draws_df(model_full_RE) %>% # 8000 draws x 299 variables
   select(starts_with("b_")) %>%       # 8000 x 18
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial100 = -900/1000)  
# effects for trial 1000
p1congruent <- post %>% 
  mutate(postdistr = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
  labs(title = "trial 1000", x = "time bin", y = "effect of congruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

p2incongruent <- post %>% 
  mutate(postdistr = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
  group_by(period_9) %>%

  ggplot(aes(x = period_9, y = postdistr)) +
  stat_lineribbon(show.legend=F) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  scale_y_continuous(breaks = c(-6,-4,-2,0,2,4,6), limits = c(-6,6)) +
  labs(x = "time bin", y = "effect of incongruent") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90))

p1congruent + p2incongruent
#(p1+p2+p3)/(p4+p5+p6)/(p7+p8+p9)/(p10+p11+p12)/(p13+p14+p15)
```

```{r save-plot-effects}
ggsave("figures/effects_knot_1.png", width = 12, height = 16, dpi = 600)

```