### 2024-09-17 Tutorial_3.Rmd
remove redundant models

---
title: "Tutorial_3"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

# Load the libraries that we will be using

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")

lapply(pkg, library, character.only = TRUE)
```

# Load the person-trial-bin data set that we saved in Tutorial 1.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
ptb_data <- ptb_data %>% filter(period > 5)

# create factor condition, with "blank" as the reference level
ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ptb_data <- ptb_data %>% 
        mutate(period_9 = period - 9,
               trial_c = (trial - 1000)/1000)

head(ptb_data,n=17)
summary(ptb_data)
# 6 subjects
# condition: blank (6401), congruent (2642), incongruent (3797)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

```{r get-data}
M3_data <- ptb_data %>% select(pid, period_9, trial_c, condition, event)
```

```{r fit-model}
model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data=M3_data, 
              family=binomial(link="cloglog"))
```

Model_M3_f took about 3 hours.

```{r save-model}
save(model_M3_f, file="Tutorial_3_Frequentist/model_M3_f.RData")
```

# Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3

```{r load-models}
model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
load("Tutorial_3_Frequentist/model_M3_f.RData")
```

```{r extract-estimates}
est <- tibble(brm = fixef(model_M3)[,1],
              glmer = fixef(model_M3_f),
              effects = names(fixef(model_M3_f))) 
```

```{r plot-estimates}
cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
shapes  <- c("s1" = 16, "s2" = 17)

p1 <- ggplot(data = est, aes(x = effects))
p1 <- p1 + geom_point(aes( y = brm,  color = "c1", shape = "s1"))
p1 <- p1 + geom_point(aes( y = glmer, color = "c2", shape = "s2")) 
p1 <- p1 + labs( x = "Effects", y = "cloglog-hazard" )
p1 <- p1 + scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer"))
p1 <- p1 + scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill="white",color="black"),
         legend.position = "top",
         axis.text.x = element_text(angle=90)) 
p1
```