### 2024-10-18 Tutorial_3b.Rmd
add tutorials 2b and 3b

---
title: "Tutorial_3b"
author: "sven panis"
date: "2024-10-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

# Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme")

lapply(pkg, library, character.only = TRUE)
```

# Load the data set for ca(t) modeling that we saved in Tutorial 1a.

```{r load-data-ca}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
print(ca_data,n=30)
summary(ca_data) # 2683 rows
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% filter(drt > 5)

# create factor condition, with "blank" as the reference level
ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ca_data <- ca_data %>% 
        mutate(period_9 = drt - 9,
               trial_c = (trial - 1000)/1000)

head(ca_data,n=17)
summary(ca_data)
# 6 subjects
# condition: blank (1303), congruent (668), incongruent (626)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit a Frequentist generalized linear mixed model (cfr. Model_M3 in Tutorial 2)

```{r get-data}
M3_ca_data <- ca_data %>% select(pid, period_9, trial_c, condition, acc)
```

```{r fit-model, eval=F}
model_M3_ca_f <- glmer(acc ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data=M3_ca_data, 
              family=binomial(link="logit"))
```

Model_M3_ca_f took about 15 minutes to fit.

```{r save-model, eval=F}
save(model_M3_ca_f, file="Tutorial_3_Frequentist/model_M3_ca_f.RData")
```