### 2024-11-07 notes/tutorial_notes.Rmd

---
title: "tutorial_notes"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is a way to provide feedback on draft versions of the tutorials. 
I can edit the tutorials directly of course (which git should track), but this is a space to provide comments.

# October 31st 2024 #

The basic workflow of the tutorials seems pretty good to me.
I can do some more fine-tuning of the code after the next revision of the manuscript,
given that some stuff will change.

## table of contents, aesthetics and apperance ##

I think we can easily improve the visual appearance of the tutorials in .html format by using a theme,
as well as their ease of navigation by setting a table of contents function.
To do so, include the following in th YAML header.

output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united

This is what we did in Andrew Wildman's psychopy tutorial. 
There are other themes, but this works for me.