### 2024-11-26 Tutorial_3a.Rmd

---
title: "Tutorial_3a"
author: "Sven Panis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

# 1. Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Load and wrangle the person-trial-bin data set.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
```

Prepare the data set.

```{r prepare-data}
# select analysis time range
ptb_data <- ptb_data %>% filter(period > 5) %>%

# create factor condition, with "blank" as the reference level
mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent"))) %>%

# center TIME (period) on bin 9, center trial on trial 1000 and rescale, 
# create variables with index coding (timebin, prime).
mutate(period_9 = period - 9,
       trial_c = (trial - 1000)/1000,
       timebin = factor(period, levels=c(6:15)),
       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
       prime = factor(prime, levels=c(1,2,3)))

head(ptb_data,n=17)
```

# 3. 3. Fit hazard model in the frequentist framework (model_M1i_f)

```{r get-data}
data_M1i_f <- ptb_data %>% select(pid, event, timebin, prime)
```

Model_M1i_f contains the same effects as Model_M1i in Tutorial 2a.

```{r fit-model, eval=F}
model_M1i_f <- glmer(event ~ 0 + timebin:prime  +
                            (0 + timebin:prime | pid),
          # control parameters, data set, and complementary log-log link function
          control = glmerControl(optimizer = c("nlminbwrap"),
                                 optCtrl = list(maxfun=10000000)), 
          data = data_M1i_f, 
          family = binomial(link="cloglog"))
```

Model_M1i_f took about 3 hours.

```{r save-M1i-f, eval=F}
save(model_M1i_f, file="Tutorial_3_Frequentist/model_M1i_f.RData")
```

```{r inspect-M1i-f}
load("Tutorial_3_Frequentist/model_M1i_f.RData")
summary(model_M1i_f)
```

# 4. Compare MLE estimates from model_M1i_f with the posterior mean estimates from model_M1i.

```{r load-models}
model_M1i <- readRDS("Tutorial_2_Bayesian/models/model_M1i.rds")
load("Tutorial_3_Frequentist/model_M1i_f.RData")
```

```{r extract-estimates}
est <- tibble(brm = fixef(model_M1i)[,c(1,3,4)],
              glmer = tidy(model_M1i_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
              effectname = names(fixef(model_M1i_f)),
              position = 1:30)

colnames(est)
```

```{r plot-estimates}
cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
shapes  <- c("s1" = 16, "s2" = 17)

p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 1.5) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 1.5, 
                 position = position_nudge(x = .2))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                show.legend = F) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F, 
                position = position_nudge(x = .2)) +
 geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 1:30,labels = est$effectname) +
 scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +

 labs( x = "Effects in model M1i", y = "cloglog-hazard" ) +
 scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
  coord_flip() #+
 #theme(panel.grid = element_blank())
         #legend.position = "top",
         #axis.text.x = element_text(size=11,angle=90),
         #axis.text.y = element_text(size=13),
         #axis.title.x = element_text(size = 14),
         #axis.title.y = element_text(size = 14),
         #legend.text=element_text(size=13)) 
p1
```

```{r save-plot-comparison, eval=F}
ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=8,dpi=800)
```