### 2024-12-05 Tutorial_1b.Rmd
added tidyverse wrangling style and reformatted code chunks

---
title: "Tutorial_1b"
author: "Sven Panis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).

# 1. Load the libraries that we will be using.

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 14, face = "bold"), 
          title = element_text(size = 14, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Read and wrangle data.

Read in the raw data from Panis and Schmidt (2016).

```{r read-raw-data}
raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", 
                  delim="\t",
                  col_names=T) 
head(raw)
```

Do some wrangling.

```{r wrangle-raw-data}
# remove practice trials
data_wr <- raw %>% 
  filter(practice==0)

# remove unnecessary columns
data_wr <- data_wr %>% 
  select(-c(ss,practice,rd,form,task,primeduration,
            blankduration,maskduration,targetduration,
            void,target_dir,corresp,resperr))

# look at trials where resp==0
data_wr %>% 
  filter(resp==0) %>% 
  print(n=51)

# remove trials with no response info if rt < 800 (800 was the response deadline in each trial):
# the 46 trials where rt >= 800 and resp==0 are trials in which the subject did not respond; these should not be deleted because they will be treated as right-censored observations in EHA!
# the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
data <- data %>% 
  filter(!(resp == 0 & rt < 800))

summary(data_wr)
# vp = Versuchsperson or subject
# bl = block number
# tr = trial number within block
# mask_type: no mask (0), relevant mask (1), irrelevant mask (2), lines mask (3)
# prime_type: no prime or blank (0), consistent prime (1), inconsistent prime (2)
# resp = no button (0), button left (1), button right (2) 
# respac = accuracy
# rt = response time
# TrialNr = trial number within experiment
```

Change variable names, and create factors.

```{r create-names-and-factors}
# change some variable names to required names (pid, trial, condition1, condition2, rt, and acc)
data_wr <- data_wr %>% 
  rename(pid = vp, condition1 = mask_type, condition2 = prime_type, 
         acc = respac, trial = TrialNr)

# pid to indicate column with unique person IDs
# trial must be a unique trial number per subject
# condition1 must be a factor indicating the levels (1,2,...) and labels of the first independent variable (here: mask_type)
# condition2 must be a factor indicating the levels (1,2,...) and labels of the second independent variable (here: prime_type)
# rt to indicate column with response times (in ms)
# acc to indicate column with accuracies (1/0)

# create factors for conditions (mask type and prime type): 
# the levels of condition1 are displayed across columns; 
# the levels of condition2 are displayed within each plot.
data_wr <- data_wr %>% 
  mutate(condition1 = condition1 + 1, 
         condition1 = factor(condition1, 
                             levels = c(1,2,3,4), 
                             labels = c("nomask","relevant","irrelevant","lines")),
         condition2 = condition2 + 1, 
         condition2 = factor(condition2, 
                             levels = c(1,2,3), 
                             labels = c("blank","congruent","incongruent")))
summary(data_wr)
```

# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)

Load the functions. Make sure you rename, and specify the same levels and labels for factors condition1 and condition2 in the function plot_eha_2IV as above.

```{r load-functions}
source("custom_functions.R")
```

Apply the functions

```{r apply-functions}
# create a data frame with two columns (pid and data); the latter contains a list of nested data frames
data_nested <- data_wr %>% 
  group_nest(pid)

# apply the functions using map() and map2()
data_final <- data_nested %>% 
  # ! user input: censoring time, and bin width
  mutate(censored  = map(data, censor, 600, 40)) %>%   
  mutate(ptb_data  = map(censored, ptb)) %>%           
  # create life tables without ca for 2 independent variables
  mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%    
  # calculate conditional accuracy for 2 independent variables
  mutate(condacc   = map(censored, calc_ca_2IV)) %>%       
  # add ca to lifetables for 2 independent variables
  mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  
  # create plots for 2 independent variables: set haz_yaxis to 1, and first_bin_shown to 2
  mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV,1,2))  
```

# 4. Extract info (life table, plots, ...)

```{r extract-info, warning=F}
# view original data (column 2 in data_final) for subject 1:
pluck(data_final, 2, 1) %>% 
  print(n=20)

# view person-trial-bin data set (column 4) for subject 1:
pluck(data_final,4,1) %>% 
  print(n=30)

# view lifetable (column 7) for a specific combination of condition levels for subject 1:
pluck(data_final, 7, 1) %>% 
  filter(condition1=="nomask" & condition2 == "blank")

# view plot (column 8) for subject 1:
pluck(data_final, 8, 1)

# view the median RTs in a tibble:
median_bin <- map(data_final$lifetable_ca, extract_median_2IV)

binsize    <- data_final$censored[[1]] %>% 
  pull(bin_width) %>% 
  max(na.rm=T)

median_rt  <- modify_depth(median_bin,1,~.x*binsize)

median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% 
  list_rbind(names_to = "pid")

median_tb %>% 
  print(n=6*12)
```

Save the plots.

```{r save-plots, warning=F, eval=F}
# save the plot for each subject
map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_2IV_for_subject", data_final$pid, "_PanisSchmidt.png"), 
     data_final$plot, ggsave, width = 12, height = 10, dpi = 800)
```