### 2024-12-09 Tutorial_2b.Rmd
update tuts

---
title: "Tutorial_2b"
author: "sven panis"
date: "2024-10-18"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data.

# 1. Load the libraries that we will be using.

```{r load-pkg, results='hide'}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel", "modelr",
         "rstan")

lapply(pkg, library, character.only = TRUE)
```

Set options.

```{r set-options, results='hide'}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()
detectCores()
```

```{r check-info, results='hide'}
packageVersion("cmdstanr")
devtools::session_info("rstan")
```

theme settings for ggplot

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Load and wrangle the person-trial data set.

```{r load-data, results='hide'}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
head(ca_data)
```

Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% 
  filter(drt > 5) %>%
  # create factor condition, with "blank" as the reference level
  mutate(condition = factor(condition, 
                            labels = c("blank", "congruent","incongruent")),
         prime = ifelse(condition=="blank",1, 
                 ifelse(condition=="congruent",2,3)),
         prime = factor(prime,levels=c(1,2,3)),
         timebin = factor(drt, levels=c(6:15)),
         period_9 = drt - 9,
         trial_c = (trial - 1000)/1000) 
head(ca_data)
```

# 3. Fit a conditional accuracy model.

We fit a single model "M1i_ca", using index coding for both TIME and condition (variables "timebin" and "prime" in ca_data).

## Prepare data for M1i_ca.

```{r data-M1i-ca}
# remove unnecessary columns before fitting a model
data_M1i_ca <- ca_data %>% 
  select(pid,acc,timebin,prime)
head(data_M1i_ca)
```

## Set up priors for M1i_ca.

```{r priors-cloglog-M1i-ca}
priors_M1i_ca <- c(
  set_prior("student_t(7.61, 0, 1.57)", class = "b"), # use flat prior on logit scale for the proportion ca 
  set_prior("normal(0, 1)", class = "sd"),                    
  set_prior("lkj(2)", class = "cor")                        
)
```

## Fit model M1i_ca.

```{r fit-model-M1i-ca, eval=F}
plan(multicore)

model_M1i_ca <- 
   brm(data = data_M1i_ca,
       family = bernoulli(link="logit"),  # use logit link for conditional accuracy
       formula = acc ~ 0 + timebin:prime  +
                        (0 + timebin:prime | pid),
       prior = priors_M1i_ca,
       chains = 4, cores = 4, iter = 3000, warmup = 1000, 
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 14),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M1i_ca")
```

Model_M1i_ca took about 11 minutes to run.

```{r inspect-M1i-ca}
model_M1i_ca <- readRDS("Tutorial_2_Bayesian/models/model_M1i_ca.rds")
fixef(model_M1i_ca)
```

# 4. Display effects for the conditional accuracy model.

## Pareto k estimates

```{r pareto}
loo(model_M1i_ca)$diagnostics %>% 
  data.frame() %>% 
  # attach the `id` values
  bind_cols(data_M1i_ca) %>% 
  mutate(id = 1:n()) %>%
  ggplot(aes(x = id, y = pareto_k)) +
  geom_point(alpha = 3/4) + 
  geom_text(data = . %>% filter(pareto_k > .2),
            aes(x = id + 2, label = id),
            size = 3, hjust = 0) +
  theme(panel.grid = element_blank())
```

## Plot posterior distributions for M1i_ca.

Wrangle the posterior draws.

```{r post-M0i}
post_ca <- as_draws_df(model_M1i_ca) %>% 
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(model_M1i_ca, robust = TRUE)
post_summary[1:30,]

post_qi_b <- post_ca %>%
  select(starts_with("b_")) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>% 
  median_qi(value) %>% 
  arrange(name)
head(post_qi_b) # 30 "fixed" effects 
```

Visualise fixed effects.

```{r plot-post-M0i-ca}
tidy_fixed_ca <- post_ca %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) %>% 
  mutate(timebin = str_sub(name,10,11),
         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
         condition = str_sub(name,17,18),
         condition = factor(str_remove(condition,"e"),
                            levels=c(1,2,3),
                            labels=c("blank","congruent","incongruent")))
head(tidy_fixed_ca)
tail(tidy_fixed_ca)

# plot
p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
  stat_halfeye(point_interval = "median_qi",
               .width = c(.80,.95),
               alpha=0.7) +
  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
       x = "time bin", 
       y = "logit-ca") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_y_continuous(limits=c(-10,10)) +
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~condition)
p_tidy_fixed_ca
```

```{r save-post, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr.png", width = 10, height = 8, dpi = 800)
```

#
```{r plot-post-M0i-ca-alt}
tidy_fixed_ca <- post_ca %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% 
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) %>% 
  mutate(timebin = str_sub(name,10,11),
         timebin = factor(str_remove(timebin,":"),levels=c(6:15)),
         condition = str_sub(name,17,18),
         condition = factor(str_remove(condition,"e"),
                            levels=c(1,2,3),
                            labels=c("blank","congruent","incongruent")))
head(tidy_fixed_ca)
tail(tidy_fixed_ca)

# plot
p_tidy_fixed_ca <- ggplot(tidy_fixed_ca, aes(x = timebin, y = value, fill=condition)) +  
  stat_halfeye(point_interval = "median_qi",
               .width = c(.80,.95),
               alpha=0.7,
               position = pd) +
  labs(title = 'Posterior distributions for population-level\neffects in Model M1i_ca',
       x = "time bin", 
       y = "logit-ca") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_y_continuous(limits=c(-10,10)) +
  theme(axis.text.x = element_text(angle=90)) #+
  #facet_wrap(~condition)
p_tidy_fixed_ca
```

```{r save-post-alt, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_postdistr_alt.png", width = 10, height = 8, dpi = 800)
```

#

## Plot predicted conditional accuracy functions

### Grand mean

```{r epred-grand}
dat_M1i_ca <- as_tibble(model_M1i_ca$data)

epreds_grand <- dat_M1i_ca %>% 
  data_grid(timebin, prime) %>% 
  add_epred_draws(model_M1i_ca, 
                  re_formula = NA) %>% 
  mutate(prime = factor(prime, 
                        levels=c(1,2,3),
                        labels=c("blank","congruent","incongruent"))) %>%
  ungroup()
```

Summarize and plot predicted hazard values.

```{r epred-grand-summary}
epreds_grand %>% 
  group_by(timebin,prime) %>% 
  median_qi()
```

```{r epred-grand-plot}
 p1 <- ggplot(epreds_grand, aes(x=timebin, y=.epred, 
                          fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") #+
    #ggtitle("Average participant")
p1
```

```{r save-epred-grand, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_grand.png", width = 10, height = 8, dpi = 800)
```

### Subject-specific

```{r epred-pid}
epreds_pid <- dat_M1i_ca %>% 
  data_grid(pid,timebin, prime) %>% 
  add_epred_draws(model_M1i_ca, 
                  re_formula = NULL) %>% 
  mutate(prime = factor(prime, 
                        levels=c(1,2,3),
                        labels=c("blank","congruent","incongruent"))) %>%
  ungroup()
```

Summarize and plot predicted hazard values.

```{r epred-pid-summary}
epreds_pid %>% 
  group_by(pid,timebin,prime) %>% 
  median_qi()
```

```{r plot-epred-pid}
p2 <- ggplot(epreds_pid, aes(x=timebin, y=.epred, 
                       fill=prime, color=prime)) +
    stat_lineribbon(point_interval="median_qi",
                    alpha = 0.5, 
                    .width = c(.8,.95)) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(0,1)) +
    labs(y = "predicted conditional accuracy") +
   # ggtitle("Each participant (N = 6)") +
    facet_wrap(~pid, ncol=3)
p2
```

```{r save-epred-pid, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_pid.png", width = 10, height = 8, dpi = 800)
```

Combine the plots.

```{r save-epred-combined, eval=F}
# combine 2 plots
p1_theme <- p1 + theme(legend.position = "none")

(p1_theme / p2) +
    plot_annotation(tag_levels = 'A') +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save plot.

```{r save-combined-pred, eval=F}
# save combined plot
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pred_combined.png", width = 10, height = 12, dpi = 800)
```

## Plot average marginal effects (AMEs)

### Grand mean AMEs

```{r grand-ame}
epreds_grand_diffs <- epreds_grand %>%
  pivot_wider(id_cols = c(timebin, .draw),
              names_from = "prime",
              values_from = ".epred") %>% #80000 x 5
  mutate(`congruent minus blank` = congruent - blank,
         `incongruent minus blank` = incongruent -  blank) %>% 
  select(-c(blank,congruent,incongruent)) %>% 
  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
               names_to = "contrast",
               values_to = "diff_ca")

epreds_grand_diffs # 160 000 rows 
```

Summarize and plot.

```{r grand-ame-summary}
table <- epreds_grand_diffs %>%
  group_by(contrast,timebin) %>%
  mean_qi(.width = c(.95)) %>%
  arrange(contrast,timebin,.width) %>%
  select(-c(.width,.point,.interval))
```

```{r save-table-grand-AMEs, eval=F}
write_csv(table, file="Tutorial_2_Bayesian/tables/ca_grand_AMEs.csv")
```

```{r grand-ame-plot}
 ggplot(epreds_grand_diffs, aes(x=timebin, y=diff_ca, 
                       fill=contrast)) +
  stat_lineribbon(alpha = 0.5, 
                  point_interval = "mean_qi",
                  .width=c(.8,.95)) +
 # stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95))+
  geom_hline(yintercept = 0, 
             color = "red", 
             lty = 3) +
  scale_y_continuous(limits=c(-1,1)) +
  scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(y = "difference in predicted ca") +
  ggtitle("Grand average marginal effect") +
  facet_wrap(~contrast)
```

```{r save-grand-AMEs, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_grand_AME.png", width = 10, height = 8, dpi = 800)
```

### Subject-specific AMEs

To calculate the subject-specific AMEs for each time bin, we create contrasts between the conditional means.

```{r epred-pid-ame}
epreds_pid_diffs <- epreds_pid %>%
  pivot_wider(id_cols = c(pid,timebin, .draw),
              names_from = "prime",
              values_from = ".epred") %>% #80000 x 5
  mutate(`congruent minus blank` = congruent - blank,
         `incongruent minus blank` = incongruent -  blank) %>% 
  select(-c(blank,congruent,incongruent)) %>% 
  pivot_longer(cols = c(`congruent minus blank`,`incongruent minus blank`),
               names_to = "contrast",
               values_to = "diff_ca") 
epreds_pid_diffs # 160 000 rows 
```

Summarize and plot the contrasts in predicted hazard values.

```{r epred-pid-ame-summary}
epreds_pid_diffs %>% 
  group_by(pid,timebin,contrast) %>% 
  mean_qi()
```

```{r epred-pid-ame-plot}
 ggplot(epreds_pid_diffs, aes(x=timebin, y=diff_ca, 
                       fill=contrast, color=contrast)) +
    stat_lineribbon(alpha = 0.5, 
                    point_interval = "mean_qi",
                    .width=c(.8,.95)) +
    #stat_halfeye(point_interval = "mean_qi",.width=c(.8,.95)) +
    geom_hline(yintercept = 0, color = "red", lty = 3) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    theme(axis.text.x = element_text(angle=90)) +
    scale_y_continuous(limits=c(-1,1)) +
    labs(y = "difference in predicted ca") +
    ggtitle("Subject-specific AMEs") +
    facet_wrap(~pid, ncol=3)
```

```{r save-pid-AMEs, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M1i_ca_pid_AMEs.png", width = 10, height = 8, dpi = 800)
```