### 2025-06-14 sims/sims.Rmd
Add code for Figure 1 in revision1

---
title: "sims"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file simulates the data used in Figures 1, 2, and 3 of the manuscript and then creates some plots.

# load the libraries that we will be using #

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "truncnorm", "faux", "ggpmisc")

lapply(pkg, library, character.only = TRUE)
```

## plot settings ##

theme settings for ggplot

```{r plot-settings}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# section 1 #

## create a baseline hazard function ##

set up and plot a user-defined discrete-time hazard function.
This hazard function takes on a reasonable shape for speeded rt tasks.
The shape serves as a reference point that other effects simulated conditions can be compared against.

```{r}
## define some variables
followup <- 1000 # length of data collection period in ms for each trial
nr_bins  <- 10
bin_width <- followup/nr_bins
bin_endpoints <- (1:nr_bins)*bin_width

## define the hazard function at each of 10 time bins
hazard_cond1 <- c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39)

## create a df
data_h <- tibble(
  hazard = hazard_cond1, 
  time = bin_endpoints
  )

## a quick plot to take a look
p1.1 <- ggplot(data_h, aes(x = time, y = hazard)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = bin_endpoints) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("endpoints of discrete time bins (ms)") +
  ylab("discrete-time hazard probability")
p1.1
```

# section 2 #

## create a function that can simulate hazard functions between conditions ##

this is to simplify as much as possible and to use it to illustrate one key message. e.g., mean conceals what time reveals

Just two conditions.

### first, let's do it step-by-step (without a function) to check the process ###

define parameters

```{r}
subj_n = 10
rep_n = 200
cond_n = 2
haz_n = 10
bin_n = 10
hazard_cond1 <- c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39)
hazard_cond2 <- hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2)
hazard <- c(hazard_cond1, hazard_cond2)
```

setup the data

```{r}
# make it reproducible
set.seed(1)

# set up data structure
dat_test <- add_random(subj = subj_n, rep = rep_n) %>% 
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
  add_within("subj", bin = 1:bin_n) %>% 
  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
         event = rbinom(n(), 1, hazf)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()

head(dat_test)
str(dat_test)
summary(dat_test)
```

take a look

data check

```{r}
dat_test %>%
  distinct(outcomel)

dat_test %>%
  tally(outcomel)

dat_test %>%
  distinct(outcomen)
```

plot

```{r}
p2.1 <- ggplot(dat_test, aes(x = bin, y = hazf, colour = condition)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bins") +
  ylab("discrete-time hazard probability")
p2.1
```

create summary data

at the pid level

```{r}
pid_dat_test <- dat_test %>% 
  group_by(subj, condition, bin) %>% 
  summarise(outcomen = mean(outcomen, na.rm=TRUE),
            n=n())
pid_dat_test
```

at the group level

```{r}
group_dat_test <- pid_dat_test %>% 
  group_by(condition, bin) %>% 
  summarise(group_outcomen = mean(outcomen, na.rm=TRUE),
            n = n(),
            group_sd = sd(outcomen, na.rm = TRUE),
            sem = (group_sd/sqrt(n)))
group_dat_test
```

plot pid and group summary data together

```{r}
p2.2 <- ggplot(group_dat_test, aes(x = bin, y = group_outcomen, 
                   colour = condition, fill=condition)) +
  geom_line(linewidth=1) + 
  geom_point(size=3) +
  geom_jitter(data=pid_dat_test, aes(y=outcomen), height=0, width = 0.1, alpha = 0.5) +
  geom_errorbar(aes(ymin = group_outcomen-sem, ymax = group_outcomen+sem),
                width=.2, linewidth=1) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x="timebin",
       y="probability")
p2.2
```

### now write a function to simulate hazard data ###

```{r}
sim_haz <- function(subj_n = 10, rep_n = 200, bin_n = 10, cond_n = 2,  # these can be changed when calling the f
               hazard_cond1 = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
               hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2),
               hazard = c(hazard_cond1, hazard_cond2),   # 
                ... # helps the function work with pmap() below
                ) {

  # set up data structure
  data <- add_random(subj = subj_n, rep = rep_n) %>% 
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
  add_within("subj", bin = 1:bin_n) %>% 
  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
         event = rbinom(n(), 1, hazf)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()

  # glimpse(data) # only use this when testing the code
}
```

Here’s a quick example of how our function works. You can change these parameters
and create some different data.

```{r}
sim_haz(subj_n = 20, rep_n = 200) # if you uncomment glimpse above,
# it will let you glimpse the data that's generated. this is useful for checking / testing code purposes.
```

test it by varying the subj_n

```{r}
data_n10 <- sim_haz(subj_n = 10, rep_n = 200)
head(data_n10)

data_n20 <- sim_haz(subj_n = 20, rep_n = 200)
```

data check

```{r}
data_n10 %>%
  distinct(subj, rep, condition)
```

a quick plot to check

```{r}
p2.3 <- ggplot(data_n10, aes(x = bin, y = hazf, colour = condition)) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bins") +
  ylab("discrete-time hazard probability")
p2.3
```

ok, this seems to work, as we would expect.

# section 3 #

## convert hazard data into rt data ##

rt at the trial level

use a uniform distribution to provide upper and lower bounds

```{r}
rt_dat_test <- data_n10 %>%
  select(-contrast, -event, -cumsum, -outcomel) %>% 
  mutate(u = bin*100,
         l = u-100,
         m = (l + u)/2) %>% 
  filter(outcomen > 0) %>% 
  group_by(subj, rep, condition, bin) %>% 
  # mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
  mutate(rt = runif(n=1, min=l, max=u))
head(rt_dat_test)
summary(rt_dat_test)
```

data check

```{r}
rt_dat_test %>% 
  group_by(condition, bin) %>% 
  summarise(min = min(rt),
            max=max(rt),
            mean=mean(rt))

rt_dat_test %>% 
  group_by(condition, bin) %>% 
  tally()
```

this looks sensible.

plot

```{r}
ggplot(rt_dat_test, aes(x=rt, colour = condition)) +
   geom_density()

rt_dat_test %>% 
  mutate(bin = factor(bin)) %>% 
  ggplot(aes(x=rt, colour = bin)) +
  geom_density() +
  facet_wrap(~condition)
```

create summary data per condition at the pid and group level for plotting mean comparisons

at the pid level

```{r}
rt_dat_pid <- rt_dat_test %>%
  group_by(subj, condition) %>% 
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # 
            sem = (sd/sqrt(n)))
rt_dat_pid
```

at the group level

```{r}
rt_dat_group <- rt_dat_test %>%
  group_by(condition) %>% 
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = length(unique(subj)), # 
            sem = (sd/sqrt(n)))
rt_dat_group
```

violin plot

```{r}
p3.1 <- ggplot(rt_dat_pid, aes(x=condition, y = mean, 
                             fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = rt_dat_group, 
             aes(y = mean), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = rt_dat_group,
                aes(y = mean, ymin = mean-sem, ymax = mean +sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = rt_dat_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   ggtitle("mean rt by condition") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   ylab("mean rt (ms)")
p3.1
```

plot together

```{r}
p3.2 <- p3.1 / p2.2
p3.2
```

# section 4 #

## create functions to automate repeat processes ##

These can then be used below to simulate, wrangle and plot the data

### create a function to summarise hazard data ###

at the pid level

```{r}
haz_pid <- function(df) {
  df %>% 
    group_by(subj, condition, bin) %>% 
    summarise(outcomen = mean(outcomen, na.rm=TRUE),
              n=n(),
              .groups = "drop")
}
```

at the group level

```{r}
haz_group <- function(df) {
  df %>% 
    group_by(condition, bin) %>% 
    summarise(group_outcomen = mean(outcomen, na.rm=TRUE),
              n = n(),
              group_sd = sd(outcomen, na.rm = TRUE),
              sem = (group_sd/sqrt(n)),
              .groups = "drop")
}
```

test them

```{r}
haz_pid_ftest <- haz_pid(data_n10)
haz_pid_ftest

haz_group_ftest <- haz_group(haz_pid_ftest)
haz_group_ftest
```

### create a function to convert hazards into rt data ###

```{r}
rt_dat <- function(df) {
  df %>%
    select(-contrast, -event, -cumsum, -outcomel) %>% 
    mutate(u = bin*100,
         l = u-100,
         m = (l + u)/2) %>% 
    filter(outcomen > 0) %>% 
    group_by(subj, rep, condition, bin) %>% 
    mutate(rt = runif(n=1, min=l, max=u)) %>% 
    ungroup()
}
```

test it

```{r}
rt_dat_ftest <- rt_dat(data_n10)
head(rt_dat_ftest)
glimpse(rt_dat_ftest)
```

### create a function to summarise rt data ###

at the pid level

```{r}
rt_pid <- function(df) {
  df %>% 
    group_by(subj, condition) %>% 
    summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # 
            sem = (sd/sqrt(n)),
            .groups = "drop")
}
```

at the group level

```{r}
rt_group <- function(df) {
  df %>% 
    group_by(condition) %>% 
    summarise(mean = mean(rt),
              sd = sd(rt), 
              n = length(unique(subj)),
              sem = (sd/sqrt(n)),
              .groups="drop")
}
```

test them

```{r}
rt_pid_ftest <- rt_pid(rt_dat_ftest)
rt_pid_ftest

rt_group_ftest <- rt_group(rt_dat_ftest)
rt_group_ftest
```

### create a function to plot the hazard data ###

hazard plot function

```{r}
haz_plot <- function(df_pid, df_group) {
  ggplot(df_group, aes(x = bin, y = group_outcomen, 
                   colour = condition, fill=condition)) +
  geom_line(linewidth=1) + 
  geom_point(size=3) +
  geom_jitter(data=df_pid, aes(y=outcomen), height=0, width = 0.1, alpha = 0.5) +
  geom_errorbar(aes(ymin = group_outcomen-sem, ymax = group_outcomen+sem),
                width=.2, linewidth=1) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x="time bin",
       y="probability")
}
```

test it

```{r}
p_haz_test <- haz_plot(haz_pid_ftest, haz_group_ftest)
p_haz_test
```

### create a function to plot the rt data ###

mean rt plot function

```{r}
rt_plot <- function(df_pid, df_group) {
   ggplot(df_pid, aes(x=condition, y = mean, 
                      fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = df_group, 
             aes(y = mean), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = df_group,
                aes(y = mean, ymin = mean-sem, ymax = mean+sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = df_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   ylab("mean rt (ms)")
}
```

test it

```{r}
p_rt_test <- rt_plot(rt_pid_ftest, rt_group_ftest)
p_rt_test
```

# section 5 #

## use the function to simulate A=B for the mean, but then some different distributions ##

simulate hazard data

```{r}
## make it reproducible
set.seed(123)

## mean rt = same, haz = same
haz1 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1))

## mean rt = same, haz = later difference, B > A
haz2 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 1.8, 1.8, 1.5, 1))

## mean rt = same, haz = later difference, A > B
haz3 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 1, 0.2, 0.2, 0.5, 1))
```

create summary data

```{r}
haz1_pid <- haz_pid(haz1) %>% 
  mutate(example = "example 1")
haz1_pid

haz1_group <- haz_group(haz1_pid) %>% 
  mutate(example = "example 1")
haz1_group

haz2_pid <- haz_pid(haz2) %>% 
  mutate(example = "example 2")
haz2_pid

haz2_group <- haz_group(haz2_pid) %>% 
  mutate(example = "example 2")
haz2_group

haz3_pid <- haz_pid(haz3) %>% 
  mutate(example = "example 3")
haz3_pid

haz3_group <- haz_group(haz3_pid) %>% 
  mutate(example = "example 3")
haz3_group
```

plot the hazard data

```{r}
p5.1 <- haz_plot(haz1_pid, haz1_group)
p5.1

p5.2 <- haz_plot(haz2_pid, haz2_group)
p5.2

p5.3 <- haz_plot(haz3_pid, haz3_group)
p5.3
```

convert to rt

```{r}
## make it reproducible
set.seed(123)

rt1 <- rt_dat(haz1)
head(rt1)

rt2 <- rt_dat(haz2)
head(rt2)

rt3 <- rt_dat(haz3)
head(rt3)
```

summarise the rt data

```{r}
rt1_pid <- rt_pid(rt1) %>%
  mutate(example = "example 1")

rt1_group <- rt_group(rt1) %>% 
  mutate(example = "example 1")

rt2_pid <- rt_pid(rt2) %>% 
  mutate(example = "example 2")

rt2_group <- rt_group(rt2) %>% 
  mutate(example = "example 2")

rt3_pid <- rt_pid(rt3) %>% 
  mutate(example = "example 3")

rt3_group <- rt_group(rt3) %>% 
  mutate(example = "example 3")
```

plot the rt data

```{r}
p5.4 <- rt_plot(rt1_pid, rt1_group)
p5.4

p5.5 <- rt_plot(rt2_pid, rt2_group)
p5.5

p5.6 <- rt_plot(rt3_pid, rt3_group)
p5.6
```

## combine the data and plot together ##

```{r}
## hazard
haz_pid_123 <- bind_rows(haz1_pid, haz2_pid, haz3_pid) %>% 
  mutate(example=factor(example))
head(haz_pid_123)

haz_group_123 <- bind_rows(haz1_group, haz2_group, haz3_group) %>% 
  mutate(example=factor(example))
head(haz_group_123)

## rt
rt_pid_123 <- bind_rows(rt1_pid, rt2_pid, rt3_pid) %>% 
  mutate(example=factor(example))
head(rt_pid_123)

rt_group_123 <- bind_rows(rt1_group, rt2_group, rt3_group) %>% 
  mutate(example=factor(example))
head(rt_group_123)
```

plot

hazard

```{r}
p5.7 <- haz_plot(haz_pid_123, haz_group_123) +
  facet_wrap(~example)
p5.7
```

rt

```{r}
p5.8 <- rt_plot(rt_pid_123, rt_group_123) +
  facet_wrap(~example)
p5.8
```

combine plots

```{r}
p5.9 <- p5.8 / p5.7
p5.9
```

# section 6 #

## use the function to simulate A>B for the mean, but then some different distributions ##

simulate hazard data

```{r}
## make it reproducible
set.seed(456)

## mean rt A>B, haz = early diff
haz4 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 3, 3, 3, 2, 1, 1, 1, 1, 1))

## mean rt A>B, haz = middle diff
haz5 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 2.5, 1.7, 1.7, 1.2, 1, 1, 1))

## mean rt A>B, haz = early diff and later reversal
haz6 <- sim_haz(subj_n = 10, rep_n = 400, 
                hazard_cond2 = hazard_cond1*c(1, 3, 3, 3, 2, 0.5, 0.4, 0.3, 0.2, 0.2))
```

create summary data

```{r}
haz4_pid <- haz_pid(haz4) %>% 
  mutate(example = "example 4")
haz4_pid

haz4_group <- haz_group(haz4_pid) %>% 
  mutate(example = "example 4")
haz4_group

haz5_pid <- haz_pid(haz5) %>% 
  mutate(example = "example 5")
haz5_pid

haz5_group <- haz_group(haz5_pid) %>% 
  mutate(example = "example 5")
haz5_group

haz6_pid <- haz_pid(haz6) %>% 
  mutate(example = "example 6")
haz6_pid

haz6_group <- haz_group(haz6_pid) %>% 
  mutate(example = "example 6")
haz6_group
```

plot the hazard data

```{r}
p6.1 <- haz_plot(haz4_pid, haz4_group)
p6.1

p6.2 <- haz_plot(haz5_pid, haz5_group)
p6.2

p6.3 <- haz_plot(haz6_pid, haz6_group)
p6.3
```

convert to rt

```{r}
## make it reproducible
set.seed(456)

rt4 <- rt_dat(haz4)
head(rt4)

rt5 <- rt_dat(haz5)
head(rt5)

rt6 <- rt_dat(haz6)
head(rt6)
```

summarise the rt data

```{r}
rt4_pid <- rt_pid(rt4) %>%
  mutate(example = "example 4")

rt4_group <- rt_group(rt4) %>% 
  mutate(example = "example 4")

rt5_pid <- rt_pid(rt5) %>% 
  mutate(example = "example 5")

rt5_group <- rt_group(rt5) %>% 
  mutate(example = "example 5")

rt6_pid <- rt_pid(rt6) %>% 
  mutate(example = "example 6")

rt6_group <- rt_group(rt6) %>% 
  mutate(example = "example 6")
```

plot the rt data

```{r}
p6.4 <- rt_plot(rt4_pid, rt4_group)
p6.4

p6.5 <- rt_plot(rt5_pid, rt5_group)
p6.5

p6.6 <- rt_plot(rt6_pid, rt6_group)
p6.6
```

## combine the data and plot together ##

```{r}
## hazard
haz_pid_456 <- bind_rows(haz4_pid, haz5_pid, haz6_pid) %>% 
  mutate(example=factor(example))
head(haz_pid_456)

haz_group_456 <- bind_rows(haz4_group, haz5_group, haz6_group) %>% 
  mutate(example=factor(example))
head(haz_group_456)

## rt
rt_pid_456 <- bind_rows(rt4_pid, rt5_pid, rt6_pid) %>% 
  mutate(example=factor(example))
head(rt_pid_456)

rt_group_456 <- bind_rows(rt4_group, rt5_group, rt6_group) %>% 
  mutate(example=factor(example))
head(rt_group_456)
```

plot

hazard

```{r}
p6.7 <- haz_plot(haz_pid_456, haz_group_456) +
  facet_wrap(~example)
p6.7
```

rt

```{r}
p6.8 <- rt_plot(rt_pid_456, rt_group_456) +
  facet_wrap(~example)
p6.8
```

combine plots

```{r}
p6.9 <- p6.8 / p6.7
p6.9
```

big plot

```{r}
# p6.10 <- (p5.9 / p6.9) +
#   plot_layout(axes = 'collect',
#               guides = 'collect')
# p6.10
```

# section 7 #

## combine all examples into one df and plot ##

### wrangle ###

```{r}
# haz
haz_pid_all <- bind_rows(haz_pid_123, haz_pid_456) %>% 
  mutate(example=factor(example))
head(haz_pid_all)

haz_group_all <- bind_rows(haz_group_123, haz_group_456) %>% 
  mutate(example=factor(example))
head(haz_group_all)

## rt
rt_pid_all <- bind_rows(rt_pid_123, rt_pid_456) %>% 
  mutate(example=factor(example))
head(rt_pid_all)

rt_group_all <- bind_rows(rt_group_123, rt_group_456) %>% 
  mutate(example=factor(example))
head(rt_group_all)
```

### plot ###

hazard

```{r}
p7.1 <- haz_plot(haz_pid_all, haz_group_all) +
  facet_wrap(~example, nrow = 1)
p7.1
```

rt

```{r}
p7.2 <- rt_plot(rt_pid_all, rt_group_all) +
  facet_wrap(~example, nrow = 1)
p7.2
```

combine plots

```{r}
p7.3 <- (p7.2 / p7.1) +
  plot_annotation(tag_levels = 'A')
p7.3

#ggsave("sims/figures/rt_haz.jpeg",
#      width = 12, height = 8, dpi=800)
```

## use patchwork to combine prior plots ##

```{r}
# p7.4 <- (p5.9 / p6.9) 
#   # plot_annotation(tag_levels = 'A')
# p7.4
# 
# ggsave("sims/figures/rt_haz_2.jpeg",
#        width = 12, height = 8, dpi=800)
```

# section 8 #

## try to inset rt within the hazard plot panels ##

create a haz plot with two rows

```{r}
p8.1 <- haz_plot(haz_pid_all, haz_group_all) +
  facet_wrap(~example, nrow = 2)
p8.1
```

re-plot with 2 columns, rather than 2 rows, and re-order the facets so that 1-3 are in col1

```{r}
p8.1b <- haz_plot(haz_pid_all, haz_group_all) +
  facet_wrap(~fct_relevel(example,'example 1','example 4','example 2','example 5',
                          'example 3', 'example 6'), ncol = 2)
p8.1b
```

create a rt plot two rows

```{r}
p8.2 <- rt_plot(rt_pid_all, rt_group_all) +
  facet_wrap(~example, nrow = 2)
p8.2
```

re-plot rt with 2 columns, rather than 2 rows

```{r}
p8.2b <- rt_plot(rt_pid_all, rt_group_all) +
  facet_wrap(~example, ncol = 2)
p8.2b
```

now let's try using ggpmisc to inset rt within the haz plots

## let's first try a simpler example ##

the rt plot should be simpler and have smaller text, if it is going to be an inset plot

let's create a couple of mean rt plots. and then facet by hazard.

### create a function for simpler rt plots ###

a simpler mean rt plot function

```{r}
rt_plot_inset <- function(df_pid, df_group) {
   ggplot(df_pid, aes(x=condition, y = mean, 
                      fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(alpha = 0.7) +
   geom_point(data = df_group, 
             aes(y = mean), size = 2, position=pd2, colour="black") +
   geom_errorbar(data = df_group,
                aes(y = mean, ymin = mean-sem, ymax = mean+sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = df_group, 
            aes(y = mean, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   scale_y_continuous(limits = c(300, 600), breaks = seq(300, 600, 100)) +
   labs(y="",
        x="") +
   theme_bw() +
   theme(legend.position = "none",
         text = element_text(face = "bold"),
         panel.background = element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank())
}
```

test it

```{r}
p_rt_inset_test <- rt_plot_inset(rt_pid_ftest, rt_group_ftest)
p_rt_inset_test
```

now create a df with the ggplot object in there, so it can be used later

```{r}
inset <- tibble(x = 0, y = 1,
             plot = list(p_rt_inset_test))
```

combine it with a hazard plot

using geom_plot

```{r}
p_haz_inset <- p2.2 +
  geom_plot(data = inset, 
            vp.width = 1/4,
            aes(x, y, label = plot))
p_haz_inset

# ggsave("sims/figures/haz_inset.jpeg",
#        width = 8, height = 8, dpi=800)
```

### now incorporate a facet ###

create six examples of the simpler rt plots

```{r}
p8.1s <- rt_plot_inset(rt1_pid, rt1_group)
p8.1s

p8.2s <- rt_plot_inset(rt2_pid, rt2_group)
p8.2s

p8.3s <- rt_plot_inset(rt3_pid, rt3_group)
p8.3s

p8.4s <- rt_plot_inset(rt4_pid, rt4_group)
p8.4s

p8.5s <- rt_plot_inset(rt5_pid, rt5_group)
p8.5s

p8.6s <- rt_plot_inset(rt6_pid, rt6_group)
p8.6s
```

create the inset data

```{r}
inset_facet <- tibble(x = rep(0, times = 6),
                      y = rep(1, times = 6), 
                      example = factor(c("example 1", "example 2", "example 3",
                                     "example 4", "example 5", "example 6")),
                      plot = list(p8.1s, p8.2s, p8.3s, p8.4s, p8.5s, p8.6s))
```

add the inset to the main plot

with two rows

```{r}
# p_haz_inset_facet <- p8.1 +
#   geom_plot(data = inset_facet,
#             vp.width = 1/3,
#             aes(x, y, label = plot))
# p_haz_inset_facet
# 
# ggsave("sims/figures/haz_inset_facet.jpeg",
#        width = 14, height = 10, dpi=1000)
```

with two columns

```{r figure-1}
p_haz_inset_facet <- p8.1b +
  geom_plot(data = inset_facet,
            vp.width = 1/3,
            aes(x, y, label = plot))
p_haz_inset_facet

ggsave("sims/figures/haz_inset_facet.jpeg",
       width = 10, height = 12, dpi=1000)
```

# section 9 #

## create a figure that illustrates hazard plus conditional accuracy ##

### create the hazard data ###

create hazard data

```{r}
## make it reproducible
set.seed(7)

## create the data
haz7 <- sim_haz(subj_n = 10, rep_n = 400,
                hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, 1, 0.5, 0.3, 0.2, 0.2, 0.2))
```

calculate summary hazard data

```{r}
haz7_pid <- haz_pid(haz7)
haz7_pid

haz7_group <- haz_group(haz7_pid)
haz7_group
```

plot hazard

```{r}
p9.1 <- haz_plot(haz7_pid, haz7_group) +
  labs(title = "hazard") +
  theme(legend.position = "none")
p9.1

## if you want to remove individual pid points for simplicity
# p9.1 <- p9.1 %>%  
#   remove_geom('point')
# p9.1
```

### convert to rt data ###

convert to rt data

```{r}
## make it reproducible
set.seed(7)

rt7 <- rt_dat(haz7)
head(rt7)
```

calculate summary rt data

```{r}
rt7_pid <- rt_pid(rt7)
rt7_pid

rt7_group <- rt_group(rt7)
rt7_group
```

simple plot rt

```{r}
# p9.2 <- rt_plot(rt7_pid, rt7_group)
# p9.2

p9.2 <- rt_plot_inset(rt7_pid, rt7_group)
p9.2
```

### inset rt within the hazard plot ###

create the inset data

```{r}
inset_hazrt <- tibble(
  x = 1,
  y = 0.9, 
  plot = list(p9.2)
)
```

and plot

```{r}
p9.3 <- p9.1 +
  geom_plot(data = inset_hazrt,
            vp.width = 1/3,
            aes(x, y, label = plot))
p9.3
```

### accuracy ###

create some data

```{r}
acc <- tibble(
  time = seq(1,10,1),
  cond1 = rep(1,length = 10),
  cond2 = c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
) %>% 
  pivot_longer(-time,
               names_to = "condition",
               values_to = "accuracy") %>% 
  mutate(condition = factor(condition))
head(acc)
```

create group level summary data

```{r}
acc_group <- acc %>% 
  group_by(condition) %>% 
  summarise(group_accuracy = mean(accuracy, na.rm = TRUE),
            n=n(),
            group_sd=sd(accuracy, na.rm = TRUE),
            sem = (group_sd/sqrt(n)),
            .groups = "drop")
acc_group
```

line plot per bin

```{r}
p9.4 <-ggplot(acc, aes(x = time, y = accuracy,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("timebin") +
  ylab("probability") +
  ggtitle("accuracy") 
p9.4
```

mean average plot

```{r}
p9.5 <- ggplot(acc, aes(x=condition, y = accuracy, 
                      fill = condition, colour = condition)) +
   geom_jitter(position=position_jitterdodge(dodge.width =1),
               alpha = 0.5) +
   geom_violin(alpha = 0.7) +
   geom_point(data = acc_group, 
             aes(y = group_accuracy), size = 2, position=pd2, colour = "black") +
   geom_errorbar(data = acc_group,
                aes(y = group_accuracy, ymin = group_accuracy-sem, ymax = group_accuracy+sem),
                width=.2, position=pd2, colour = "black") +
   geom_line(data = acc_group, 
            aes(y = group_accuracy, group=1),
            position=pd2,  colour = "black") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   scale_y_continuous(limits = c(0, 1), breaks = seq(0,1,0.25)) +
   labs(y="",
        x="") +
   theme_bw() +
   theme(legend.position = "none",
         text = element_text(face = "bold"),
         panel.background = element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank())
p9.5
```

inset mean acc within the acc plot

create the inset data

```{r}
inset_acc <- tibble(
  x = 1,
  y = 0.9, 
  plot = list(p9.5)
)
```

and plot

```{r}
p9.6 <- p9.4 +
  geom_plot(data = inset_acc,
            vp.width = 1/3,
            aes(x, y, label = plot))
p9.6
```

## plot together ##

```{r}
p9.7 <- (p9.3 / p9.6) +
  plot_layout(guides = 'collect',
              axes = 'collect')
p9.7

## without inset
# ggsave("sims/figures/haz_acc.jpeg",
#        width = 8, height = 8, dpi=800)

## with inset
# ggsave("sims/figures/haz_acc_inset.jpeg",
#        width = 8, height = 8, dpi=800)
```

It doesn't make sense to create the inset with this data because the bins (1:10) are just to illustrate the point rather than being tied to actual RTs, like in a real study where bin 3 may represent 250-300ms say.

# section 10 #

## create a figure that illustrates hazard and survivor functions ##

create the data

```{r}
data_hs <- tibble(
  timebin = seq(1,10,1),
  hazard = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
  survivor = c(1, 0.97, 0.91, 0.79, 0.61, 0.39, 0.22, 0.08, 0.04, 0.02)
) %>% 
  pivot_longer(-timebin,
               names_to = "dv",
               values_to = "probability") %>% 
  mutate(dv = factor(dv))
head(data_hs)
```

plot

```{r}
p9.7 <-ggplot(data_hs, aes(x = timebin, y = probability,
                         colour = dv)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("hazard and survivor functions") 
p9.7

#ggsave("sims/figures/haz_surv.jpeg",
#       width = 7, height = 5, dpi=800)
```

# Section 11 (Figure 2): Simulate hazard data for 1 subject, 1 condition, 200 trials, and plot empirical hazard and survivor functions with error bars

## Simulate hazard data based on user-defined hazard function

```{r}
## make it reproducible
set.seed(7)

## apply function sim_haz
data_single <- sim_haz(subj_n = 1, rep_n = 200) %>%
               filter(condition == "cond1") %>%
               mutate(period = bin,
                      trial = str_sub(rep, start=6, end=6)) %>%
               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
head(data_single, n=20)
```

## set up life table

```{r}
lifetable <- data_single %>%
    mutate(event = str_c("event", event)) %>%
    group_by(condition,period) %>% 
    count(event) %>% 
    ungroup() %>% 
    pivot_wider(names_from = event,
                values_from = n) %>% 
    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
           event1 = ifelse(is.na(event1),0,event1),
           risk_set = event0 + event1) %>% # define the risk set
    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
           pmass = event1/max(risk_set),
           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
    group_by(condition) %>%
    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
    ungroup() 
```

## wrangle life table for plotting

```{r}
minimallt <- lifetable %>% select(period,condition, hazard, se_haz, survival, se_surv) %>%
  pivot_longer(cols = c(hazard,survival),
               names_to = "dv",
               values_to = "probability") %>%
  mutate(se = ifelse(dv=="hazard", se_haz, se_surv))
```

## plot hazard and survivor functions

```{r}
p11.1 <-ggplot(minimallt, aes(x = period, y = probability,
                         colour = dv)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=probability-se, ymax=probability+se), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("hazard and survivor functions") +
  xlab("time bin")
p11.1

ggsave("sims/figures/haz_surv_single.jpeg",
       width = 7, height = 5, dpi=800)
```

# Section 12 (Figure 3): plot of hazard and ca(t) for 2 conditions with error bars

## generate data set

```{r}
## make it reproducible
set.seed(8)

## apply function sim_haz
ntrials <- 1000

data_single2 <- sim_haz(subj_n = 1, rep_n = ntrials) %>%
               mutate(period = bin,
                      trial = as.integer(str_sub(rep, start=4, end=7)),
                      condition = factor(condition, labels=c("congruent","incongruent"))) %>%
               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep"))  %>%
               mutate(acc = ifelse(period<6 & condition=="incongruent",0,1),
                      acc = ifelse(period==6 & condition=="incongruent" & trial < ntrials/2,0,acc)) 
head(data_single2, n=20)
summary(data_single2)
```

## set up life table

```{r}
lifetable2 <- data_single2 %>%
    mutate(event = str_c("event", event)) %>%
    group_by(condition,period) %>% 
    count(event) %>% 
    ungroup() %>% 
    pivot_wider(names_from = event,
                values_from = n) %>% 
    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
           event1 = ifelse(is.na(event1),0,event1),
           risk_set = event0 + event1) %>% # define the risk set
    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
           pmass = event1/max(risk_set),
           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
    group_by(condition) %>%
    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
    ungroup() 
```

## get ca(t)

```{r}
ca <- data_single2 %>% 
    group_by(condition,period) %>%
    summarize(ca = mean(acc) %>% round(digits = 2),
              n = n(),
              .groups = 'drop') %>%
    ungroup() %>%
    mutate(se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) 
ca
```

## plot hazard functions

```{r}
p12.1 <-ggplot(lifetable2, aes(x = period, y = hazard,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("hazard")  +
  xlab("time bin") + ylab("probability") + guides(color=guide_legend(title="prime type"))
p12.1
```

## plot ca

```{r}
p12.2 <-ggplot(ca, aes(x = period, y = ca,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bin") +
  ylab("probability") +
  ggtitle("conditional accuracy") + 
  guides(color=guide_legend(title="prime type"))
p12.2
```

## Combine plots

```{r}
p12.3 <- (p12.1 / p12.2) +
  plot_layout(guides = 'collect',
              axes = 'collect')
p12.3

#ggsave("sims/figures/haz_acc_single.jpeg",
#       width = 7, height = 6, dpi=800)

```

# section 13: REVISION 1: Create New Figure 1

In Figure 1 of the revision, we want to display for 1 subject the hazard and ca functions for 2 conditions, and the corresponding four means.

Steps:

## 1. Generate discrete time-to-event data for 1 subject, 2 conditions, 100 trials, and 10 time bins;

Adjust function:

```{r}
sim_haz <- function(subj_n = 10, rep_n = 200, bin_n = 10, cond_n = 2,  # these can be changed when calling the f
               hazard_cond1 = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
               hazard_cond2 = hazard_cond1*c(1, 1, 1, 1, .5, 0.3, 0.7, 1, 1,1),
               hazard = c(hazard_cond1, hazard_cond2),   # 
                ... # helps the function work with pmap() below
                ) {

  # set up data structure
  data <- add_random(subj = subj_n, rep = rep_n) %>% 
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) %>%
  add_within("subj", bin = 1:bin_n) %>% 
  mutate(hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
         event = rbinom(n(), 1, hazf)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()
}
```

Generate data:

```{r}
## make it reproducible
set.seed(7)

n_trials=200
## apply function sim_haz
data_fig1_hazard <- sim_haz(subj_n = 1, rep_n = n_trials) %>%
               mutate(subj = factor(subj),
                      trial = as.integer(str_sub(rep, start=4, end=6)),
                      acc = ifelse(bin<5 & condition=="cond2",0,1),
                      acc = ifelse(bin==5 & trial < 100,0,acc)) %>%
               select(-c("contrast","cumsum","outcomel","outcomen","rep"))

head(data_fig1_hazard, n=20)
summary(data_fig1_hazard)
```

Set up life table for generated data:

```{r}
lifetable_fig1 <- data_fig1_hazard %>%
    mutate(event = str_c("event", event)) %>%
    group_by(condition,bin) %>% 
    count(event) %>% 
    ungroup() %>% 
    pivot_wider(names_from = event,
                values_from = n) %>% 
    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
           event1 = ifelse(is.na(event1),0,event1),
           risk_set = event0 + event1) %>% # define the risk set
    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
           pmass = event1/max(risk_set),
           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
    group_by(condition) %>%
    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
    ungroup() 
```

Calaculate conditional accuracy:

```{r}
ca <- data_fig1_hazard %>% 
    filter(event == 1) %>%
    group_by(condition,bin) %>%
    summarize(ca = mean(acc) %>% round(digits = 2),
              n = n(),
              .groups = 'drop') %>%
    ungroup() %>%
    mutate(se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) 
ca
```

## 2. Plot hazard and conditional accuracy

```{r}
p13.1 <-ggplot(lifetable_fig1, aes(x = bin, y = hazard,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10, limits = c(1,10),labels = c(1:10)*100) +
  scale_y_continuous(limits = c(0,1)) +
 # ggtitle("hazard and conditional accuracy functions") +
  xlab("time bin endpoint") +
  theme(axis.text.x = element_text(angle = 90))
p13.1

#ggsave("sims/figures/haz_surv_single.jpeg",
#       width = 7, height = 5, dpi=800)
```

```{r}
p13.2 <-ggplot(ca, aes(x = bin, y = ca,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10, limits=c(1,10),labels = c(1:10)*100) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("time bin endpoint") +
    ylab("conditional accuracy") +
   theme(axis.text.x = element_text(angle = 90))

 # ggtitle("conditional accuracy functions") 
#+ 
#  guides(color=guide_legend(title="prime type"))
p13.2
```

Combine plots

```{r}
p13.3 <- (p13.1 / p13.2) +
  plot_layout(guides = 'collect',
              axes = 'collect')
p13.3
```

## 3. Create RT data (assuming uniform distribution within a bin) and calculate means

use a uniform distribution to provide upper and lower bounds

```{r}
rt_dat_fig1 <- data_fig1_hazard %>%
  filter(event==1) %>% 
  mutate(u = bin*100,
         l = u-100,
         m = (l + u)/2) %>% 
  group_by(subj, trial, condition, bin) %>% 
  # mutate(rt = rtruncnorm(n=1, mean=m, sd=25, a=l, b=u))
  mutate(rt = runif(n=1, min=l, max=u))

head(rt_dat_fig1)
summary(rt_dat_fig1)
```

create summary RT data per condition at the pid level for plotting mean comparisons

```{r}
rt_dat_fig1_pid <- rt_dat_fig1 %>%
  group_by(condition) %>% 
  summarise(mean = mean(rt),
            sd = sd(rt),
            n = n(), # 
            sem = (sd/sqrt(n)))
rt_dat_fig1_pid
```

create summary ACC data per condition at the pid level for plotting mean comparisons

```{r}
acc_dat_fig1_pid <- rt_dat_fig1 %>%
  group_by(condition) %>% 
  summarise(mean = mean(acc),
            sd = sd(acc),
            n = n(), # 
            sem = (sd/sqrt(n)))
acc_dat_fig1_pid
```

## 4. plot rt + accuracy

```{r}
p13.4 <- ggplot(rt_dat_fig1_pid, aes(x = condition, y = mean, fill=condition, color=condition)) +
  geom_col(alpha = 0.9, width = 0.6) +
  geom_errorbar(
    aes(ymin = mean - sem, ymax = mean + sem),
    width = 0.2,
    linewidth = 0.8,
    color = "black"
  ) +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
 #  ggtitle("mean RT by condition") +
   scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, 100)) +
   ylab("mean RT (ms)")
p13.4
```

```{r}
p13.5 <- ggplot(acc_dat_fig1_pid, aes(x = condition, y = mean, fill=condition, color=condition)) +
  geom_col(alpha = 0.9, width = 0.6) +
  geom_errorbar(
    aes(ymin = mean - sem, ymax = mean + sem),
    width = 0.2,
    linewidth = 0.8,
    color = "black"
  ) +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
 #  ggtitle("mean accuracy by condition") +
   scale_y_continuous(limits = c(0, 1)) +
   ylab("accuracy")
p13.5
```

plot together

```{r}
p13.6 <- (p13.4 / p13.5) +
  plot_layout(guides = 'collect',
              axes = 'collect')
p13.6

plot_grid(p13.6, p13.3,
  nrow = 1,
  labels = c("A","B"),
  label_size = 22
)

```

```{r}
ggsave("sims/figures/Fig1_revision1.jpeg",
       width = 10, height = 10, dpi=800)

```