### 2025-06-24 manuscript/reviews/ResponseReview1.Rmd
update Response to Reviewers

---
title: "Review 1: Author responses"
author: "Panis & Ramsey"
date: "24.06.2025"
output: 
  pdf_document:
      keep_tex: true
header-includes:
  - \usepackage{framed}
  - \usepackage{xcolor}
  - \let\oldquote\quote
  - \let\endoldquote\endquote
  - \renewenvironment{quote}{\begin{leftbar}}{\end{leftbar}}
---

```{r setup, include=FALSE}
# By default, the code chunks are hidden for brevity. Set the echo argument to TRUE for blocks you want to share the code in the output for. 
knitr::opts_chunk$set(echo = F)
```

```{r packages and data, include=F }

```

# Editor

> 15-Mar-2025
>
> Dear Dr. Panis:
>
> Thank you for submitting your Tutorial  entitled "Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows" . I have now received input from three experts in the field, and you will find the reviews below this message and attached. As you will see, all three reviewers believe your paper can make a strong contribution to the literature. Prior to considering the reviewers' input, I read your paper to form an independent opinion. I share the positive sentiments about this work and invite you to revise and resubmit your manuscript for further consideration.

> The reviews are through and thoughtful, and I will not reiterate all the details here. For my part, I would like your revision to focus on accessibility. Your paper is a bit challenging-- it's long and a little tough to get through. What can make it easier for less familiar readers? Perhaps more of a basic set up earlier? A glossary defining specific terms (e.g., what is an event? how is it defines? censoring? hazards vs. survival curves, conditional probabilities, etc.)? Also, clarify when readers need to shift from frequentist to Bayesian model... and why? The main consideration here is that wish to make the paper much more accessible-- from start to finish-- for readers who are novices and entirely unfamiliar with this area. This can be done any number of ways, and R3 also includes some suggestions for making these reviews. 
I encourage you to have the revised paper read and pre-reviewed by colleagues who are novices in this area with the explicit goal of making the paper accessible to them-- this will help you hit the more for our broad readership.

> If you choose to submit a revision, please include include a letter detailing your point-by-point responses to each reviewer comment and indicating how you changed the manuscript to address them. The revised manuscript may undergo further peer review. We ask that you submit your revision within three months. Please let us know if you will not be able to meet this deadline.

> Sincerely,

Response: Thank you for allowing us to submit a revision of our Tutorial (AMPPS-24-0232).

We tried to focus on accessibility in the revision of our Tutorial in several ways. 
First, we replaced Figure 1 with a simpler Figure that directly conveys our message, and deleted Figures 2 and 3.
Second, we shortened the Introduction and Discussion sections.
Third, we added a section in the supplementary material (section A) in which we visualize the different types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception).
We also mention the unique contribution of our work, we removed redundancies, and tried to improve the clarity of the paper.

# Reviewer: 1

> Comments to the Author
> It was a pleasure to review this manuscript as someone who has spent years applying survival analysis methods to psychological data. This is an ambitious manuscript and I believe it achieves its goals. The key contributions of this manuscript from my perspective are the creation of custom functions for data wrangling, and the section on planning studies/simulations for power analysis (of course, in addition to the introduction to this category of methods and the practical tutorials provided. I have some comments on this work that I believe could strengthen this manuscript’s contribution.

> General comments:

> The Introduction is really nicely framed in terms of providing compelling information for why this class of statistical models would be useful in some specific areas of psychology (i.e., reaction time data). Given that there are already a couple of tutorial-types of articles introducing specific applications of survival analysis in developmental psychology (references below, both regarding the use of these methods for behavioural observations of emotional expressions in a multilevel framework for recurring events), it would be helpful to further frame the unique contribution that this work adds. The authors may also consider further emphasizing the specific theoretical contributions that using these methods may lead to in the areas of reaction time data etc.

Response: In the revised ms. we frame the unique contribution that this works adds on pages 5-6:
"...we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which
provide worked examples of the key data processing and Bayesian multilevel regression modelling steps."

> Lougheed, J. P., Benson, L., Ram, N., & Cole, P. M. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53–65. https://doi.org/10.1037/dev0000619

> Stoolmiller, M. (2016). An introduction to using multivariate multilevel survival analysis to study coercive family process. In T. J. Dishion & J. J. Snyder (Eds.), The Oxford handbook of coercive relationship dynamics (pp. 363–378). Oxford University Press.

> Stoolmiller, M., & Snyder, J. (2006). Modeling heterogeneity in social interaction processes using multilevel survival analysis. Psychological Methods, 11(2), 164–177. https://doi.org/10.1037/1082-989X.11.2.164

Response: We added these references to the ms.

> Similarly, it would be helpful to make it clear that this manuscript pertains to estimating single-event occurrences rather than recurring events (which are described in the Lougheed and Stoolmiller tutorials mentioned above)

Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant.

> Specific comments:

> The authors use several different terms to refer to time-to-event models throughout the manuscript. The title refers to EHA, and the running head refers to Hazard analysis. Survival analysis is also a common term in Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller & Snyder, 2006). Disambiguating the various terms early in the manuscript and then constantly using one term throughout the manuscript would be helpful.

Response: We disambiguate the various term in the revised ms. on page 7, and use one term throughout.

> Page 5: it would be helpful to refer to the same time unit in reference to Figure 1, which shows time bins along the x-axis but the examples discussed in the manuscript refer to units of milliseconds.

Response: Figure 1 has been replaced.

> Page 7: The authors state that they are not aware of EHA tutorials for psychological time to event data. Previoulsy mentioned tutorial articles by Lougheed and Stoolmiller could be mentioned here— these are demonstrating applications to behavioural observation data within psychology.

Response: We added these 3 references to the revised ms.

> Page 8: I’m curious about the decision to focus on discrete time rather than continuous time models. It would help the reader to explain the decision either way, or refer the reader to Lougheed et al., 2019 who discuss this distinction and choices that researchers may make regarding how to handle decisions around time given data formats. I don’t work with reaction time data, but it seems to me that many experiments in this area would collect data in continuous time intervals, and it’s not clear to me what the benefit of discretizing into time bins is. Now that I’ve read the entire manuscript, I see that some information is presented on this on Page 50, but it would be helpful to make this information known to the reader earlier in the manuscript.

Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:

"...the definition of hazard and the type of models employed depend on whether one is using
continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
more experience using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
treated as a continuous variable. However, continuous forms of EHA require much more
data to estimate the continuous-time hazard (rate) function well. Thus, by trading
a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
for dealing with typical psychological RT data sets for which there are less than ~200 trials
per condition per experiment. Moreover, as indicated by Allison (2010),
learning discrete-time EHA methods first will help in learning continuous-time methods, so
it seems like a good starting point."

> Page 8-9, excellent and accessible explanations of the survival function in the discrete time framework.

Response: We moved the explanation of the survival function to the Supplementary Material in the revised ms.

> Page 11: In my experience, researchers in psychology are generally not familiar with the concept of right censored data. This term is first used on page 11 with no explanation— a brief, even parenthetical, definition would be helpful. For example, the parenthetical definition on page 48 could be provided on page 11.

Response: We introduce the definition of right-censoring on page 10 of the revised ms.

> Taken together, this is an excellent manuscript with the potential to further bring survival analysis methods into Psychology, where time-to-event and right-censored cases abound, yet very few people seem to consider these issues.

# Reviewer: 2

> Comments to the Author
> Please see attached file for comments.

> Reviewer report for AMPPS-24-0232: “Event History Analysis for psychological time-to-
event data: A tutorial in R with examples in Bayesian and frequentist workflows”
Summary: The authors present a tutorial paper on the analysis of discrete time-to-event
outcomes. They motivate their tutorials with a description of psychological experiments
that produce time-to-event data and explain how simply summarizing event-time
outcomes using means results in a loss of potentially important information. They then
explain how more information can be gleaned from event-time data if they are
summarized using event history analysis methods. They focus specifically on the
analysis of discrete time-to-event outcomes. They provide code for conducting the
analyses in the tutorials on their Github. Overall, I like the idea of this paper and
strongly agree that using methods appropriate for analyzing time-to-event data, rather
than simply summarizing event-time outcomes using means, is very important and
currently underused. While I like the premise of the paper and think that such a paper
could be a valuable contribution to the literature, I do think the paper has room for
improvement in a number of areas (see comments below). The writing is approachable
(which is advantageous for a tutorial paper) but some text in the paper was repetitive
and could be a bit more polished; additional editing with a focus on clarity and brevity
would be beneficial.

Response: We removed repetitive text, and try to focus on clarity and brevity in the revised ms.

> Major comments:
1. Some discussion on the choice of bin width is needed. Since all of the example
time-to-event data is collected in continuous time, the analyst must discretize it
by defining bins. Defining bins is a key step in the analysis pipeline and so
should not be overlooked. It seems like results—and potentially conclusions
drawn—may depend on the choice of bin width. What factors should the analyst
consider when defining the width of the bins? Does the number of bins matter?
Or is the number of events per bin important? What about other factors? Should
analysts (or potentially these authors too) consider multiple bin widths when
conducting an analysis? An example comparing results when bin width varies
could be both interesting and informative.

Response: We discuss the issue of defining bins on page 11 of the revised ms.:

"Third, the width of each time bin will need to be determined. For instance, in Figure
1B we chose 100ms in an arbitrary manner. In reality, however, bin width will need to be
set by considering a number of factors simultaneously. The optimal bin width will depend
on (a) the length of the observation period in each trial, (b) the rarity of event occurrence,
(c) the number of repeated measures (or trials) per condition per participant, and (d) the
shape of the hazard function. Finding an appropriate bin width in a given user case before
fitting models will require testing a number of options, when calculating and plotting the
descriptive statistics (see section 3.1). The goal is to find the smallest bin width that is
supported by the amount of data available. Based on our experience, a bin width of 50 ms
is a good starting value when the number of repeated measures is 100 or less. Too small
bin widths will result in erratic hazard functions as many bins will have no events, and
thus hazard estimates of zero. Interestingly, the time bins do not need to have the same
width. For example, Panis (2020) used larger bins towards the end of the observation
period, as fewer events occurred there."

> 2. In the discussion, the authors mention that they choose to present a tutorial on
discrete time analysis methods because they believe these are easier to
understand. My personal impression is that awareness of continuous time
analysis methods is slightly higher and so researchers may be more familiar with
methods such as Cox models, even if they have not used them themselves;
however, this could vary by specific sub-field. I agree with the authors that the
interpretation of the hazard is easier in the discrete time setting than in the
continuous time setting, but I do think that the distinction between the
interpretation of the hazard in continuous vs discrete time should be discussed in
the introduction. When I first read the introduction, I was confused as to why the
focus was on discrete time methods, when the outcome measures collected in
the motivating data were collected in continuous time. I would suggest either of
the following: (a) provide concrete examples of when truly discrete timeoutcomes are collected to strengthen the motivation of the paper and justification
for use of discrete time analysis methods or (b) keep the motivating examples as-
is and include a short comparison of discrete vs. continuous time analysis
methods to help explain your reasoning for choosing to focus on discrete time
analysis methods.

Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:

"...the definition of hazard and the type of models employed depend on whether one is using
continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
more experience using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
treated as a continuous variable. However, continuous forms of EHA require much more
data to estimate the continuous-time hazard (rate) function well. Thus, by trading
a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
for dealing with typical psychological RT data sets for which there are less than ~200 trials
per condition per experiment. Moreover, as indicated by Allison (2010),
learning discrete-time EHA methods first will help in learning continuous-time methods, so
it seems like a good starting point."

> 3. Censoring is an important part of time-to-event data and so more description of
censoring—and how it is (not) accounted for in “orthodox” analysis methods vs.
discrete time-to-event analysis methods would be useful, beyond the already-
included mention of censoring in Section 5.3.

Response: We introduce the definition of right-censoring on page 10 of the revised ms.

> 4. Section 1.2 is a bit long. Please shorten this section to improve the clarity of the
text and reduce redundancy. There is also overlap between text in Section 1.2
and the start of Section 4.

Response: We shortened this section and removed the overlap.

> 5. Section 2.3.2. I like the idea of making a connection with experimental design,
but am having trouble seeing how the discussion in this section is specific to EHA
and how these comments do not apply also to the “orthodox method”. I think this
section could be improved by providing more specific distinctions between
experimental design implications for EHA vs. the “orthodox method”.

Response: We created a new Figure 1 that makes the intended point, and removed section 2.3.2.

> 6. Line 517: What are model weights? Please explain briefly for the reader,
including how the interpretation differs for loo and waic. Also, in the provided
reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
than WAIC, so could you please include a definition of or reference for dWAIC in
your context? Lastly, when reporting the model weights in the example in the
text, I’d suggest including more than 2 digits so that it is clear if the reported
weights are actually 0 and 1 or if rounding has just made them appear so.

Response: TODO...

We explain model weights on page x, include a reference for dWAIC, and report the weights with 3 digits.

> 7. Tutorials 3a and 3b: Rather than simply stating that the frequentist models did
not converge and resulted in singular fits, it would be helpful to provide more
discussion. Do you have an idea why the models did not converge? Since this
is a tutorial paper, more explanation is needed. These frequentist examples
don’t add much to the paper in their current form, so I’d either suggest either (a)
providing some suggestions to the reader of alternative approaches or specific
strategies that could be used to define a model that will converge or (b) removing
these examples from the main text and directing the reader to the supplement
instead.

Response: We removed the frequentist examples from the main text.

> Minor comments:
> 1. Is “mean-average comparison” standard terminology? This phrase seems
redundant to me, since means and averages are the same thing.

Response: We removed this term.

> 2. Figure 1: In the caption, please state what the small error bars in the inset plots
are showing.

Response: The old Figure 1 has been replaced.

> 3. Please consider updating all plots so that they are understandable when viewed
in black and white. In addition to using different colors, I’d suggest distinguishing
the conditions using shapes (e.g., circles vs. triangles) and different line type
(e.g., solid vs. dashed lines).

Response: TODO

> 4. Figure 2 caption:a. Please provide a reference for the statement beginning, “Because the
survival function…”.

Response: Figure 2 has been removed. We only discuss the survival function in the Supplemental Material (section B) of the revised ms.

> b. I am confused by the sentence beginning: “For example, the high hazard
of …”. Could you please provide some additional explanation or a
reference for more details on this interpretation?

Response: We removed this section.

> 5. Lines 156-166: Generally, survival probabilities are described as probabilities of
remaining event-free up until a certain point, rather than the probability of an
event occurring in the future. Some clarity in wording could be used here.

Response: We follow this definition in section B of the Supplemental Material.

> 6. Please confirm that you have defined all acronyms, abbreviations, and functions
at their first use in the text. E.g., please define WAIC, LOO, P(t), ca(t), SOA, CrI,
among others.

Response: We define all acronyms on their first mention in the revised ms.

> 7. Section 3.1 could be condensed in the main paper (additional details could be
included in the supplementary material if important). Also, please double check
your definition of functional programming.

Response: Section 3 has been removed in the revised ms.

> 8. Table 2: It could be helpful to show an example of a trial where an event never
occurs, to help the reader connect what is written in the text (line 325-327) to
what appears in the table.

Response: We added data visualizations in section A of the Supplemental Material.

> 9. Figure 4: Why is there no bin 40?

Response: TODO

> 10. Line 423: The phrase “early responses” is confusing as it implies temporal
ordering of the responses (i.e., early vs. late trials). I’d suggest using “fast
response” or “rapid responses”, or some other phrase along those lines.

Response: TODO

> 11. Is the discussion of reference and index coding (lines 457 – 464) needed? In the
introduction, the authors state that knowledge of regression is assumed so I am
not sure this is necessary to include, but I defer to the authors’ judgment.

Response: Index coding is not used a lot in experimental psychology so we included a short discussion.

> 12. Figure 6: I only see one width of credible interval for each line on this plot—
please clarify if you are plotting 80% or 95% credible intervals.

Response: TODO -> correction needed.

> 13. Line 536: To me, “subject-specific” and “marginal” mean the opposite things.
That is, subject-specific is an effect that is conditional on an individual while a
“marginal” effect has been marginalized/averaged over individuals. Could you
please clarify your use of these terms here?

Response: TODO.

> 14. Lines 584-587: Since this is a tutorial paper, I’d suggest including explanations
for why you make changes (a), (b), and (c).

Response: TODO.

> 15. Line 659: What is a “fully varying effects structure”? Please define/explain.

Response: TODO.

> 16. Line 832-833: Please provide a reference for the statement, “However, they
require much more data…”.

Response: TODO.

> 17. Github code: Instruction 2 of the README states “open the
reproducible_workflow.Rproj” – I can’t find this R project, should it be the
“Tutorial_EHA.Rproj” instead?

Response: Yes this has been corrected now.

# Reviewer: 3

> Comments to the Author
This tutorial guides readers through how to wrangle cognitive reaction time data and address research questions pertaining to temporal dynamics of cognition using event history analysis (EHA) and speed/accuracy tradeoff analysis (SAT). Models are illustrated for hazard functions, and conditional accuracy functions. A nice feature of this study is that each model is presented in both the Bayesian and frequentist frameworks. Having both frameworks provides side-by-side comparisons such as when the Bayesian framework might be preferred (e.g., when models do not converge in the Frequentist framework, which is common when modeling multilevel data). The tutorial also includes a section on simulation and power analysis for planning experiments. I believe this tutorial could eventually be a useful contribution to fellow psychological scientists. However, in its current state, it was difficult to follow/ make my way through (despite existing expertise in EHA), and also difficult to understand what the main contributions were above and beyond existing works. Additionally, it was not clear why the discrete time framework was chosen. The temporal granularity of reaction time data seems like it would be better suited for a continuous time framework. Altogether, the tutorial covers important concepts, but it would benefit from greater accuracy and clearer definitions of key terms. Carefully removing redundancies, getting to the tutorial section much quicker (currently not until pg. 17), and being more precise with explanations of technical jargon would enhance its clarity and accuracy.

Response: We discuss the main contributions above and beyond existing works on pages 5-6 of the revised ms. We address the issue of treating time as discrete or continuous on pages 7-8 of the revised ms. We removed redundancies, get to the tutorial section quicker (page 11) and included more precise explanations of technical jargon.

> <b>Code for the tutorials:</b> I went to the github page and found a folder for each of the 4 tutorials, which included sub-folders for data, figures, and tables. It took a minute for me to realize that the actual R code for each tutorial was in the main folder, not within each sub-folder. This may also be confusing to readers. Additionally, each of the 4 tutorials, along with their sub-tutorials are provided in separate .Rmd files. There were several download steps needed to get all the information. To me, this made the user experience tedious, and made the tutorials seem disjointed from one another. The authors might consider a more comprehensive single Rmd file that is neatly organized with an interactive table of contents (in the knitted html file). To me, this would make the experience much more enjoyable and useful for readers. Given that this is a tutorial, it would also be nice if the the RMarkdowns were somewhat more standalone e.g., by including a brief description of what a life table is and how to interpret the plot.

Response: We explain the location of the files in the README file. A single download step is possible by either cloning the project, or downloading the ZIP file.
We included a brief description of a life table and how to interpret the plot in the R Markdown tutorial files.

> <b>Contribution of this tutorial:</b>
It wasn’t clear to me as a reader what the unique contribution is for this tutorial and what gaps in the literature it fills beyond existing texts and tutorials. For example, is this a tutorial that is meant to specifically focus on experimental psychology cognitive task data – and that this type of data has special characteristics/associated research questions that necessitate their own stand-alone tutorial? If so, it may be helpful to give more detail on what these specific characteristics tend to be that may make it different from the many other types of psychology data. Or, is this a tutorial on discrete time survival analysis that is meant to be broadly applicable to psychological scientists, and the cognitive reaction time data is just for illustrative purposes. If so, then much less emphasis should be placed on the specific data type and more emphasis should be placed more broadly on types of psychological data/ research questions this method could be applied to. I think the distinction makes a big difference in the entire framing of the tutorial, and choosing one or the other will make the tutorial content much easier for readers to absorb.

Response: This tutorial is meant to specifically focus on experimental psychology RT data and this type of data has special characteristics that necessitate their own stand-alone tutorial.
Therefore, we added visualizations of the types of time-to-event data that are obtained in typical RT tasks (detection, discrimination, bistable perception) in section A of the Supplemental Material.

> On page 7, it states that there are not tutorials aimed at psychological time-to-event data. Yet, there are at least a few existing empirical article examples with accompanying tutorials in psychology (listed below), and arguably some of the popular textbooks are also geared toward psychological scientists (e.g., Singer & Willett). Similarly, many of the textbooks contain code to accompany their text. Regardless, even if those are not counted, I still think it will be beneficial for readers if the authors can make a clear case for what is unique about psychological time-to-event data that necessitates a stand alone tutorial/ what gaps in the existing textbooks/empirical articles/tutorials the present tutorial serves to fill.

Response: In the revised ms. we frame the unique contribution that this works adds on pages 5-6:
"...we are not aware of any tutorials that are aimed specifically at psychological RT (+ accuracy) data, and which
provide worked examples of the key data processing and Bayesian multilevel regression modelling steps."

> Lougheed, J. P., Benson, L., Cole, P. M., & Ram, N. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53.

> Elmer, T., van Duijn, M. A., Ram, N., & Bringmann, L. F. (2023). Modeling categorical time-to-event data: The example of social interaction dynamics captured with event-contingent experience sampling methods. Psychological Methods.

> Mills, M. (2011). The fundamentals of survival and event history analysis. Introducing Survival Analysis and Event History Analysis. London: SAGE Publications, 1-17.

Response: We included these references in the revised ms.

> <b>Organization of the tutorial:</b>The current version of the tutorial is much longer than a typical AMPPStutorial. There are a few possibilities for dealing with this. First, it may be that the tutorial could be split into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own. Another possibility is that the focus could be primarily on the Bayesian models, and then leave the frequentist models for either a supplement, or to only live in the RMarkdown. This possibility may also be beneficial due to the model convergence issues with at least one of the frequentist models (which is common with multilevel data and why many psychological scientists are moving to the Bayesian framework). A third possibility is that the overall tutorial text could be greatly streamlined, with major cuts to both the introduction and the discussion sections.

Response: We focus primarily on the Bayesian models in the revised ms., and keep Tutorial 4.  We also made major cuts to both the introduction and the discussion sections to shorten the ms.

> <b>Clarity and accuracy: </b>
-There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided. For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.

Response: We introduce the definition of right-censoring on page 10 of the revised ms. We discuss what it means to be at risk in section A of the Supplemental Material.

> Additionally, many key details, such as the EHA equations are relegated to supplementary materials. In my opinion, any key information that makes the tutorial more accessible and precise should not be in supplementary materials.

Response: We placed the equations for h(t) and ca(t) in Figure 1 of the revised ms.

> There are many abbreviations that were difficult to follow. EHA is arguably standard, but SAT (for speed/accuracy trade-off) is less familiar, and may therefore benefit from not being an abbreviation. Similarly referring to hazard function as h(t), conditional accuracy function as c(t) was less clear than just saying hazard function and conditional accuracy function.

Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in main body of the revised ms.

> It would be helpful if the research questions that can be addressed by the methods presented are clearly presented in each section. I think that currently they are getting lost in the tutorial.

Response: TODO.

> There are several opportunities to describe the method more precisely. For example, the discrete time hazard function gives a ‘conditional’ probability (pg. 9).

Response: Indeed, this was a big error on our part. Thank you for pointing this out.

> Additionally, it may be helpful to use more standard notations such as h(t_ij) to indicate the conditional probability that the event occurs in bin j for individual i, given that the individual had not already experienced the event in a previous time period (lines 159-160, pg. 9).

Response: TODO. We add indexes for individual and trial.

>  Similarly, the survival probability is the probability that the event occurs after bin j, *given that the individual lhas survived up until bin j*.

Response: TODO.

> Throughout the tutorial, reference to individuals in the notation is missing.

Response: TODO. We added reference to individuals and trials (repeated measures) in the notation.

> Another inaccuracy is that in the discrete time framework, it is not an instantaneous likelihood (line 161, pg. 9).

Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood. We replaced the term "likelihood" with "risk".

>  <b>Introduction:</b>
-Section 1.1 – I had a difficult time understanding this section as the opening to the tutorial, and the last paragraph seemed out of place/abrupt. felt too vague to be understandable as the opening to the tutorial.

Response: We rewrote this section.

> Given that the speed-accuracy tradeoff analysis is described as central to this tutorial, it may benefit from being introduced earlier than pg. 7.

Response: We introduce conditional accuracy functions already in Figure 1 of the revised ms.

> It seems that comparing and contrasting Bayesian and frequentist methods may be outside the scope of a tutorial (pg. 7)

Response: We agree and focus now on the Bayesian models in the revised ms.

> As mentioned above, tutorial 4 – while related – could potentially be its own stand-alone tutorial.

Response: We decided to keep tutorial 4.

> Censoring is a key component of survival analysis that is missing from the introduction

Response: We introduce the definition of right-censoring on page 10 of the revised ms.

> It might be useful to indicate early on the distinctions between multilevel data, single events, and recurring events – and that this particular tutorial is only focused on single events.

Response: On page 7 of the revised ms., we mention that we measure single-event occurrences that are repeatedly measured in each participant. In section A of the Supplemental Material we mention that this Tutorial does not cover recurrent events.

> It’s not clear why a discrete time framework is used given the granular temporal resolution of reaction time data – this would also mean that the data would not need to be separated into bins. Most textbooks (e.g., Allison’s) often indicate discrete time methods are common for temporal scales such as months, years, or decades (not milliseconds). Clearer justification for the discrete time framework and separating the data into bins would be helpful.

Response: We address our choice for discrete-time methods on pages 7-8 of the revised ms.:

"...the definition of hazard and the type of models employed depend on whether one is using
continuous or discrete time units. As a lab, and mainly for practical reasons, we have much
more experience using discrete-time EHA, and that is the approach that we describe and
focus on in this paper. This choice may seem counter-intuitive, given that RT is typically
treated as a continuous variable. However, continuous forms of EHA require much more
data to estimate the continuous-time hazard (rate) function well. Thus, by trading
a bit of temporal resolution for a lower number of trials, discrete-time methods seem ideal
for dealing with typical psychological RT data sets for which there are less than ~200 trials
per condition per experiment. Moreover, as indicated by Allison (2010),
learning discrete-time EHA methods first will help in learning continuous-time methods, so
it seems like a good starting point."

> Clearer description of the data would be helpful for readers (as well as only having this description in one place in the tutorial to be more concise). For example, on pg. 10 it is indicated that 200 trials of 1 experimental condition are simulated. Can the event occur at any time in those 200 trials? If the event occurs in trial 150, do the remaining 50 trials still occur, or does the time series end for that person? How are those extra trials after an event occurred handled in the data wrangling and analysis. Can events occur more than once?

Response: We provided a clearer description of the data.

> Section 2.2 (pg. 10, lines 168-169), if describing what ‘statisticians and mathematical psychologists say’, it would be helpful to provide supporting references.

Response: TODO. We provide supporting references on page x of the revised ms.

> The numbered list on pg. 11 contained jargon statements without explanations. Suggest to simplify and keep the focus on the data you are using and specific research questions you are using survival analysis as a tool to address.

Response: We removed the numbered list in the revised ms.

> In section 2.3, suggest to lead with the data and research questions of interest and then show how features of survival analysis make it a good method to examine the research questions. I also felt that up until now, the tutorial has been very confusing and hard to follow. It could be that starting the tutorial at section 2.3 (with a few extra details filled in), may be an alternative.

Response: Section 2.3 has been rewritten. The new Figure 1 in the tutorial directly focuses  on the comparison between analysing means vs. event occurrence.

> Again, section 2.3.1 (the majority of pg. 12) was unclear and difficult to follow for someone who already has knowledge of survival analysis, but not much familiarity with cognitive process/reaction time data.

Response: Section 2.3.1 has been removed.

> From what is written, I did not have a clear sense of how temporal states of cognitive processes and theory development could occur (pg. 14, lines 231-29). These ideas and the research questions are scattered in many places throughout the tutorial, and are not very clear/straightforward. It may be helpful to locate them in just one place in the tutorial and provide a clearer description to guide readers. Similarly, the statement on pg. 14 (lines 238-239) felt rather strong without any references, and also out of scope for a tutorial.

Response: We explain the idea of temporal states of cognitive processes now on page 5 of the revised ms. and illustrate it in Figure 1.

> It wasn’t clear how section 3.1 added value to the tutorial

Response: Section 3 has been removed.

> Section 3.2 – it would be helpful if there was clarification in the text on when it would not be possible to include varying intercepts and slopes (i.e., at least clarifying that random intercepts are a requirement, but sometimes random slopes are removed based on model convergence issues – however, this also begs the question of whether a model should even be presented if it does not converge when it includes random slopes…)

Response: TODO.

> Additionally in section 3.2, the text on pg. 16 is vague and does not give much detail to readers.

Response: Section 3 has been removed.

> Overall, the introduction spans across 16 pages, which feels long even for a standard empirical article, let alone a tutorial. As currently written, I think many readers will get lost before even getting to the tutorials section.

Response: We agree, and restructured and shortened the introduction.

> <b>Tutorials section:</b>
-Rather than only using h(t), and ca(t), suggest to use hazard function and conditional accuracy function along with their mathematical notation. Without having presented any equations, using terms such as h(t) and ca(t) seems odd.

Response: We try to limit the number of abbreviations used in the main body of the revised ms.

> Similarly, P(t) is never defined for readers.

Response: P(t) was and still is defined only in the Supplementary Material (section B).

> Section 4.1.1. – it would be helpful to clarify that the risk set also pertains to individuals (not just bins)

Response: TODO.

> The connection between trials and bins could be made clearer (this is in reference to the text on pg. 18, but I think the clarity would need to come earlier on in the tutorial when describing the data)

Response: This connection should be clear now based on section A of the Supplemental Material, and based on the text on pages 7 and 8.

> Comprehension of the text on pg. 18 would be enhanced if both right and left censoring had been clearly introduced previously. Although left censoring is probably unlikely in the reaction time data used in the tutorial – if it is the case that this tutorial is meant for broader data types in psychology, left censoring would certainly be relevant to many other use cases.

Response: We introduce various types of censoring in section A of the Supplemental Material.

> Section 4.1.2 was a clearer data description than had been presented up until that point in the tutorial. Again, it is suggested that the manuscript/tutorial could be dramatically reduced in length if information was more concisely but descriptively written in a single place (e.g., data descriptions).

Response: We provide the information on data descriptions in a single place in the revised ms.

> Suggest to be more precise with the language on pg. 21 (line 355) – which data are nested within participants? The trial data? The binned data?

Response: TODO.

> pg. 22 (line 358) the way that censoring is described here is confusing.

Response: TODO. We agree and adjusted the description.

> pg. 22 discusses warning messages and lets readers know these can be ignored but does not give details as to why. Since this is a didactic tutorial, it may be helpful to have at least some explanation as to why the warning can be ignored.

Response: On line 362 it was stated that "the warning messages are generated because some bins have no hazard and ca(t) estimates, and no error bars. They can thus safely be ignored".

We removed these messages from the paper and mention them now only in the tutorial itself.

> pg. 22 – since this is a tutorial, it might also be helpful to unpack in more detail the section about identifying individuals who may be guessing etc – especially since these are very important considerations.

Response: TODO.

> section 4.3.1 – (line 451) it would be helpful if more details is given on how to set the analysis time window. What counts as “enough data”? Given that this is a tutorial, these types of detail are critical for readers to be able to implement the method in their own work

Response: TODO.

> section 4.3.1 (line 453) – more detail is needed on the cloglog link function and how it differs from the more commonly known about logit link.

Response: A comparison between the cloglog and logit link functions is provided in section C of the Supplemental Material.

> section 4.3.1 (line 456) – what goes into the decision about whether to treat time as a categorical or continuous predictor?

Response: TODO.

> section 4.3.3. – it might be helpful to give a brief indicator up front to readers about the rationale/purpose of fitting both M0i and M1i, especially given that no interpretation is given for M0i. Similarly, it would be helpful to give a bit more detail on the purpose of the comparison of M0i and M1i (given that this is a didactic tutorial).

Response: TODO.

> section 4.3.6. – same thing here – since this is a tutorial, at least brief details on what the purpose of this section is would be helpful. It might also be helpful here to define what is meant by credible interval so that more novice readers don’t confuse it with a confidence interval.

Response: TODO.

> The term SOA is used on pg. 34, but had never previously been defined or described.

Response: Stimulus-onset-asynchrony (SOA) is a well-known concept in experimental psychology.

> The text on pg. 35 (lines 567-574) seems too in-depth for a tutorial

Response: We decided to keep this example, as it can help readers with interpreting effects in hazard and conditional accuracy functions.

> pg. 36 – it would be helpful to give more detail on how to interpret the logit-ca scale

Response: We decided not to include more detail because logistic regression is very familiar to experimental psychologists.

> section 4.7 – as previously mentioned, it is possible that tutorial 4 could be a stand alone tutorial. If it’s not though, more details are needed up front to tell readers what the purpose is and why this is useful.

Response: We decided to keep tutorial 4, and discuss its usefulness on page 45 of the revised ms.

> <b>Discussion section:</b> The Discussion section would benefit from being shortened (AMPPS guidelines indicate discussions should be a brief summary of their contents rather than a general discussion).

Response: We shortened the discussion section.

> is ‘how long’ a component of survival analysis – or is it more accurate to say ‘whether and when psychological states occur’?

Response: It is more accurate to say ‘whether and when’ psychological states occur. However, based on those results, we can deduce information about how long a psychological state lasts within a trial.

Nevertheless, we replaced this with "how effects evolve with increasing waiting time".

> pg. 47 (lines 753-761) – I had a difficult time understanding this section

Response: TODO.

> section 5.5 – unclear that the multi-state model is actually an example where a multi-state model would be relevant… (the conceptual example given seems like it would be more linked to either a hidden Markov model or recurring event survival model)

Response: We removed section 5.5.

>  <b>Figures:</b>
-Figure 1 is difficult to follow. It is unclear how the means are being calculated. For example, it seems like in examples 2 and 3, the means between conditions should be different. I was also confused in example 4 because the difference in probability between the two conditions in time bin 2 seems tiny. I was also confused because it still looked like there was a big difference in probability between the two conditions at bin 5. Was bin 6 by chance what was meant instead of bin 5? It also still looked like there were differences later on such as in bin 7. Example 5 also left me confused. It looked to me like the effect started at bin 4 and ended at bin 8. Clarification (and perhaps simplifying) this figure would be useful.

Response: We created a new Figure 1.

> The description for figure 2 is unclear, particularly with respect to the contrasting knowledge that can be gained from the survival function and hazard function

Response: Figure 2 has been removed.