### 2025-07-13 Tutorial_1a.Rmd
remove bin brackets and readme.html; update Figs 1,2 and Table 3

---
title: "Tutorial 1a"
author: "Sven Panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In Tutorial 1a we wrangle raw data from Experiment 1 of @panisWhatShapingRT2016, and illustrate how to calculate descriptive statistics (functions of discrete time). We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.

A life table includes for each time bin indexed by t, the risk set (i.e., the number of trials that are event-free at the start of the bin), the number of observed events, and the estimates of the discrete-time hazard probability h(t) = P(RT = t | RT >= t), survival probability S(t) = P(Rt > t), probability mass P(t) = P(RT = t), possibly the conditional accuracy ca(t) = P(correct | RT = t), and their estimated standard errors (se).

To set up a life table, the user has to supply both a width in ms for the time bins, and a censoring time, the last time point for which the trials are considered under investigation (a cutoff time, or timeout; the response deadline if one is used). The goal is to find a good bin width and censoring time (both in ms) to portray the effects of interest. Note that all response times must be positive.

# 1. Load the libraries that we will be using.

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Read and wrangle data.

Read in the raw data file from @panisWhatShapingRT2016 - can be skipped. This file can be found on ResearchGate.

```{r read-raw-data}
raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", 
                  delim="\t",
                  col_names=T)
```

Do some wrangling to get a simpler data set - can be skipped.

```{r wrangle-raw-data}
# remove practice trials
data <- raw %>% 
  filter(practice==0) # 9240 experimental trials: 6 subj x 70 blocks x 22 trials-per-block

# remove unnecessary columns
data <- data %>% 
  select(-c(ss,practice,rd,form,task,primeduration,
            blankduration,maskduration,targetduration,
            void,target_dir,corresp,resperr))

# look at trials where resp==0
data %>% 
  filter(resp==0) %>% 
  print(n=51)

# remove trials with no response info if rt < 800 (800 was the response deadline in each trial):
# the 46 trials where rt >= 800 and resp==0 are trials in which the subject did not respond; these should not be deleted because they will be treated as right-censored observations in EHA!
# the remaining 5 trials of subject 1 should be removed (for some technical reason the resp (1 or 2) was not recorded; perhaps (s)he pressed both buttons at the same time...)
data <- data %>% 
  filter(!(resp == 0 & rt < 800))

data %>%
  distinct(resp)

# keep only the no-mask trials
data <- data %>% 
  filter(mask_type == 0) %>%
  select(-mask_type)

summary(data)
# vp = Versuchsperson or subject
# bl = block number
# tr = trial number within block
# prime_type: no prime or blank (0), consistent prime (1), inconsistent prime (2)
# resp = no button (0), button left (1), button right (2) 
# respac = accuracy
# rt = response time
# TrialNr = trial number within experiment
```

```{r save-data-file}
# save wrangled data file
write_csv(data, "Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
```

Read in simpler data file, change variable names, and create factors.

```{r read-wrangled-data}
data_wr <- read_csv("Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
head(data_wr)

# change some variable names to required names: pid, trial, condition, rt, and acc.
data_wr <- data_wr %>% 
  rename(pid = vp, condition = prime_type, acc = respac, trial = TrialNr) %>%

# pid to indicate column with unique person IDs
# trial must be a unique trial number per subject
# condition must be a factor indicating the levels (1,2,...) and labels of the single independent variable 
# rt to indicate column with response times (in ms)
# acc to indicate column with accuracies (1/0)

# create factor for condition (prime type)
  mutate(condition = condition + 1,
         condition = factor(condition,
                            levels = c(1,2,3), 
                            labels = c("blank","congruent","incongruent")))
summary(data_wr)
```

# 3. Apply functional programming to set up life tables and plot the discrete-time functions h(t), S(t), ca(t), and P(t)

Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 119-122).

```{r load-functions}
source("custom_functions.R")
```

Apply the custom functions

```{r apply-functions}
# create a dataframe with two columns (pid and data); 
# the latter contains a list of nested data frames
data_nested <- data_wr %>% 
  group_nest(pid)

# apply the functions using map() and map2()
data_final <- data_nested %>% 
    # ! user input: censoring time, and bin width
    mutate(censored = map(data, censor, 600, 40)) %>%
    # create person-trial-bin data set
    mutate(ptb_data = map(censored, ptb)) %>%
    # create life tables without ca
    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
    # calculate ca
    mutate(condacc = map(censored, calc_ca)) %>%     
    # create life tables with ca
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
    # create plots, and set haz_yaxis to 1, and first_bin_shown to 2
    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha, 1, 1))  
    # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
    #            Recommended to set this to 1 when plotting the data of individual subjects.
    # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
    #            you can set first_bin_shown to another value, such as 5 (i.e., do not show the 
    #            first 4 bins in the plots).
```

Save the calculations.

```{r save-descr, eval=F}
save(data_final, file="Tutorial_1_descriptive_stats/data/descr_stats_tut1a.Rdata")
```

# 4. Extract info (life table, plots, ...) and create files for fitting hazard and conditional accuracy models

```{r extract-info, warning=F}
# view original data (column 2) for subject 1:
pluck(data_final, 2, 1)

# view person-trial-bin data set (column 4) for subject 1:
pluck(data_final,4,1) %>% 
  print(n=30)

# view lifetable (column 7) for each condition for subject 1
pluck(data_final, 7, 1) %>% 
  filter(condition=="blank") 
pluck(data_final, 7, 1) %>% 
  filter(condition=="congruent")
pluck(data_final, 7, 1) %>% 
  filter(condition=="incongruent")

# view plot (column 8) for subject 1. 
pluck(data_final, 8, 6)

# view the median RTs in a tibble (median_tb)
median_bin <- map(data_final$lifetable_ca, extract_median)

binsize    <- data_final$censored[[1]] %>% 
  pull(bin_width) %>% 
  max(na.rm=T)

median_rt  <- modify_depth(median_bin,1,~.x*binsize)

median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% 
  list_rbind(names_to = "pid")
median_tb
```

Note that warning messages are generated when plucking a plot, like this:
"1: Removed 2 rows containing missing values or values outside the scale range".
These warning messages are generated because some bins have no hazard and ca(t) estimates, and also no error bars. They can thus safely be ignored.

Now save the plots.

```{r save-plots, eval=F}
# save the plot for each subject
map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), 
     data_final$plot, ggsave, width = 8, height = 8, dpi = 800)
```

Create a life table.

```{r create-single-lifetable}
# get the binsize in ms
binsize    <- data_final$censored[[1]] %>% 
  pull(bin_width) %>% 
  max(na.rm=T)

# extract the lifetable (column 7 in data_final) for condition 1 (blank) for subject 6
lifetable_blank_s6 <- data_final[[7]][[6]] %>% 
  filter(condition=="blank") %>% 
  mutate(bin = period*binsize,
         t = period,
         events = event1) %>%
  select(bin,t,risk_set,events,hazard,se_haz,
         survival,se_surv,ca,se_ca,pmass,se_pmass)

# get riskset for first bin
riskset <- lifetable_blank_s6 %>% 
  select(risk_set) %>% 
  slice(1) %>% 
  pull()

# create a row for time point 0
row0 <- c(bin=0,t=0,risk_set=riskset,events=NA,hazard=NA,se_haz=NA,
          survival=1,se_surv=0,ca=NA,se_ca=NA,pmass=0,se_pmass=0)

# add new row
lifetable_blank_s6_complete <- lifetable_blank_s6 %>%
  add_row(!!! row0) %>%
  arrange(bin)

# change names
names(lifetable_blank_s6_complete) <- c("bin","index t", 
                                        "RS","#events", "h(t)", "se[h(t)])", "S(t)", "se[S(t)]", "ca(t)", "se[ca(t)]", "P(t)", "se[P(t)]")

lifetable_blank_s6_complete
```

```{r save-single-lifetable, eval=F}
write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_blank_s6.csv")
```

Create input file for fitting hazard models.

```{r create-file-hazard-modeling}
# select the person-trial-bin data set for each subject, unnest them, and select relevant variables.
input_modeling_hazard <- data_final %>% 
  select(pid, ptb_data) %>%  
  unnest(ptb_data) %>%
  select(pid,bl,tr,trial,condition,period,event)
summary(input_modeling_hazard)
```

```{r save-file-hazard-modeling, eval=F}
write_csv(input_modeling_hazard, file="Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
```

Create input file for fitting ca(t) models.

```{r create-file-ca-modeling}
# select the censored data set for each subject, unnest them, discard trials that are 
# censored (RT > 600 or no response detected before the deadline), and select relevant variables. 
input_modeling_ca <- data_final %>%
  select(pid, censored) %>% 
  unnest(censored) %>%
  filter(right_censored == 0) %>% # only keep uncensored trials
  select(pid,bl,tr,trial,condition,acc,drt)
```

```{r save-file-ca-modeling, eval=F}
write_csv(input_modeling_ca,   file="Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
```

# 5. Optional: set up life tables and make one plot for data aggregated across participants

Note: This is only recommended when all subjects show the same qualitative patterns.

```{r agg-data}
# create single subject number, and unique trial number
data_agg <- data_wr %>% 
  mutate(oldtrial=trial,
         oldpid=pid,
         trial = 1:n(),
         pid = 1)
```

```{r agg-descr}
# apply functions
data_agg_nested <- data_agg %>% 
  group_nest(pid)

data_agg_final <- data_agg_nested %>% 
    mutate(censored  = map(data, censor, 600, 40)) %>%
    mutate(ptb_data  = map(censored, ptb)) %>%          
    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
    mutate(condacc   = map(censored, calc_ca)) %>%      
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
    # set haz_yaxis to .7, first_bin_shown to 3, aggregated_data to TRUE, and Nsubj to 6 
    mutate(plot = map2(.x=lifetable_ca, .y=pid, plot_eha, .7, 3, T, 6)) 
  # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
  #            Recommended to set this to 1 when plotting the data of individual subjects.
  # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
  #.           you can set first_bin_shown to another value, such as 3 (i.e., do not show the 
  #            first3 bins in the plots).
  # aggregated_data: When plotting data aggregated across individuals, set to TRUE. 
  #            This only controls the title of the plots.
  # Nsubj: The number of individuals before aggregation. Only used to set the title of the plot.
```

```{r agg-plot, warning=F}
# view plot
pluck(data_agg_final, 8, 1)
```

# 6. References.