### 2025-07-14 Tutorial_3b.Rmd

---
title: "Tutorial 3b"
author: "sven panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of @panisWhatShapingRT2016 using only the no-mask trials (prime = blank, congruent, or incongruent). Note that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure [@eager2017mixedeffectsmodelsterrible].

# 1. Load the libraries that we will be using.

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed", "rstan")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Load and wrangle the person-trial data set.

```{r load-data-ca}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
```

Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% filter(drt > 5) %>%

# create factor condition, with "blank" as the reference level; center TIME; center and rescale trial; create index-coded variables timebin and prime
mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")),
       period_9 = drt - 9,
       trial_c = (trial - 1000)/1000,
       timebin = factor(drt, levels=c(6:15)),
       prime = ifelse(condition=="blank",1, ifelse(condition=="congruent",2,3)),
       prime = factor(prime, levels=c(1,2,3)))

head(ca_data)
```

# 3. Fit a conditional accuracy model in the frequentist framework (model M1i_ca_f).

```{r get-data}
data_M1i_ca_f <- ca_data %>% select(pid, acc, timebin, prime)
```

Model_M1i_ca_f contains the same effects as Model_M1i_ca in Tutorial 2b.

```{r fit-model, eval=F}
model_M1i_ca_f <- glmer(acc ~ 0 + timebin:prime  +
                        (0 + timebin:prime | pid),

              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data = data_M1i_ca_f, 
              family = binomial(link="logit"))
```

Model_M1i_ca_f took about 90 minutes to fit on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).

```{r save-model, eval=F}
save(model_M1i_ca_f, file="Tutorial_3_Frequentist/model_M1i_ca_f.RData")
```

```{r effects-M1i-ca-f}
load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
fixef(model_M1i_ca_f)
```

# 4. Compare MLE estimates from model_M1i_ca_f with the posterior mean estimates from model_M1i_ca.

```{r load-models}
model_M1i_ca <- readRDS("Tutorial_2_Bayesian/models/model_M1i_ca.rds")
load("Tutorial_3_Frequentist/model_M1i_ca_f.RData")
```

```{r extract-estimates}
est <- tibble(brm = fixef(model_M1i_ca)[,c(1,3,4)],
              glmer = tidy(model_M1i_ca_f, 
                           effects = c("fixed"), 
                           conf.int=T, 
                           conf.level = 0.95)[,c(2,3,7,8)], 
              effectname = names(fixef(model_M1i_ca_f)),
              position = rep(6:15,3),
              prime = factor(rep(1:3, each=10),
                             labels=c("blank","congruent","incongruent")) )

est %>% print(n=30)      
```

```{r plot-estimates}
# check hex codes for color paletter
mypal <- brewer.pal(8,"Dark2")

# specify colors and shapes
cols    <- c( "c1" = mypal[1], "c2" = mypal[2])
shapes  <- c("s1" = 16, "s2" = 17)

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 2.4) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 2.4, 
                 position = position_nudge(x = .4))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                linewidth = 1.2,
                show.legend = F) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F,
                linewidth = 1.2,
                position = position_nudge(x = .4)) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +
 labs(x = "time bin (upper bound in ms)", y = "logit-ca") +
 scale_color_manual(name = "Model", 
                    breaks = c("c1", "c2"), 
                    values = cols,
                    labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                    breaks = c("s1", "s2"),
                    values = shapes,
                    labels = c("brm", "glmer")) +
 theme(panel.grid.minor = element_blank(),
       axis.text.x = element_text(angle=90)) +
facet_wrap(~prime)
p1
```

```{r save-plot-comparison, eval=F}
ggsave("Tutorial_3_Frequentist/Comparison_ca.png", height=4, width=10,dpi=800)
```

# 5. References.